### 2024-03-24 outputs/paper.qmd
create graphs and models, as well as cleaning and downloading scripts

---
title: "Untitled"
author: "Boxuan Yi"
format: pdf
---

```{r}
library(arrow)
library(readr)
library(here)
library(ggplot2)
library(janitor)
library(rstanarm)
library(dplyr)
library(modelsummary)
library(broom)
library(knitr)
```

```{r}
# clean_data <- read_parquet(file = "inputs/data/cleaned_mental_health.parquet")
clean_data <- read_csv(
    file = here("inputs/data/cleaned_mental_health.csv"),
    show_col_types = FALSE
  )
```

```{r}
age_levels <- c("Under 27", "27 to 34", "35 to 39", "40 and above")
income_levels <- c("Less than $20,000", "$20,000 - $39,999", "$40,000 - $69,999", 
                   "$70,000 - $99,999", "$100,000 - $124,999", "$125,000 - $149,999",
                   "$150,000 - $174,999", "$175,000 - $199,999", "$200,000+")
month_levels <- c("Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
clean_data$age_group <- factor(clean_data$age_group, levels = age_levels)
clean_data$household_income <- factor(clean_data$household_income, levels = income_levels)
clean_data$month <- factor(clean_data$month, levels = month_levels)
```

```{r}
# Pastel1
clean_data |>
  ggplot(mapping = aes(x = household_income, fill = age_group)) +
  geom_bar(position = "dodge2") + 
  theme(legend.position = "bottom") +
  scale_fill_brewer(palette = "Set2") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  labs(x = "Household Income in 2019", y = "Number of Observations", fill = "Age Group") 
```
```{r}
clean_data |> group_by(household_income) |> 
  summarise(average_depression = mean(EPDS), average_anxiety = mean(PROMIS)) |> 
  rename("Household Income in 2019" = household_income, 
         "Average Level of Depression" = average_depression, 
         "Average Level of Anxiety" = average_anxiety) |>
  kable()
```

```{r}
clean_data |> group_by(age_group) |> 
  summarise(average_depression = mean(EPDS), average_anxiety = mean(PROMIS)) |> 
  rename("Age Group" = age_group, 
         "Average Level of Depression" = average_depression, 
         "Average Level of Anxiety" = average_anxiety) |>
  kable()
```

```{r}
clean_data |>
  ggplot(aes(x = threaten_life, y = EPDS, color = language)) +
  geom_smooth(method = "lm") +
  theme_classic() + 
  labs(x="Perceived Level of Danger", y = "Level of Prenatal Depression") + ylim(0,1)

clean_data |>
  ggplot(aes(x = threaten_baby_danger, y = EPDS, color = language)) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x="Perceived Danger to Unborn Baby's Life", y = "Level of Prenatal Depression") + ylim(0,1)

clean_data |>
  ggplot(aes(x = threaten_life, y = PROMIS, color = language)) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x="Perceived Level of Danger", y = "Level of Prenatal Anxiety") + ylim(0,1)

clean_data |>
  ggplot(aes(x = threaten_baby_danger, y = PROMIS, color = language)) +
  geom_smooth(method = "lm") +
  theme_classic() +
  labs(x="Perceived Danger to Unborn Baby's Life", y = "Level of Prenatal Anxiety") + ylim(0,1)
```

```{r}
clean_data |>
  ggplot(aes(x = month, y = EPDS)) +
  geom_boxplot() +
  theme_classic()

clean_data |>
  ggplot(aes(x = month, y = PROMIS)) +
  geom_boxplot() +
  theme_classic()
```

# Model
```{r}
depression_model <- readRDS(file = here::here("model/depression.rds"))
```

```{r}
anxiety_model <- readRDS(file = here::here("model/anxiety.rds"))
```

# Results
```{r}
modelsummary(
  list(
    "Level of Depression" = depression_model))
```

```{r}
modelsummary(
  list(
    "Level of Anxiety" = anxiety_model))
```

```{r}
modelplot(depression_model, conf_level = 0.95) +
  labs(x = "95% Level of Depression Interval")
```

```{r}
modelplot(anxiety_model, conf_level = 0.95) +
  labs(x = "95% Level of Anxiety Interval")
```

# Discussions
```{r}
clean_data |>
  ggplot(aes(x = EPDS, y = birth_weight, color = delivery_mode)) +
  geom_smooth(method = "lm")+
  theme_minimal() + labs(x = "The Level of Prenatal Depression", y = "Birth Weight (grams)")

clean_data |>
  ggplot(aes(x = PROMIS, y = birth_weight, color = delivery_mode)) +
  geom_smooth(method = "lm")+
  theme_minimal() +
  labs(x = "The Level of Prenatal Anxiety", y = "Birth Weight (grams)")
```

# Appendix
```{r}
clean_data |>
  ggplot(aes(x = threaten_baby_danger, y = threaten_baby_harm)) +
  geom_jitter(color='darkgray', alpha=0.4) +
  geom_smooth() +
  theme_classic() 
```

```{r}
clean_data |>
  ggplot(aes(x = household_income, fill = language)) +
  stat_count(position = "dodge") +
  theme_minimal() +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

```

### 2024-03-24 scripts/01-download.R

@@ -1,0 +1,22 @@
+#### Preamble ####
+# Purpose: Download and Read in Prenatal Mental Health Data from National Library of Medicine
+# Author: Boxuan Yi
+# Email: boxuan.yi@mail.utoronto.ca
+# Date: 24 March 2024
+# Prerequisites: Know where to download the dataset
+
+# The dataset was downloaded manually from https://osf.io/ha5dp/, 
+# and uploaded as inputs/data/OSFData_Upload_2023_Mar30.csv
+
+# install.packages('rstanarm')
+# install.packages('modelsummary')
+# install.packages('ggplot2')
+# install.packages('dplyr')
+# install.packages('readr')
+# install.packages('janitor')
+# install.packages('arrow')
+# install.packages('here')
+# install.packages('broom')
+# install.packages('knitr')
+
+raw_data <- read.csv("inputs/data/OSFData_Upload_2023_Mar30.csv")

### 2024-03-24 scripts/02-clean.R

@@ -1,0 +1,55 @@
+#### Preamble ####
+# Purpose: Clean the Dataset from National Library of Medicine
+# Author: Boxuan Yi
+# Email: boxuan.yi@mail.utoronto.ca
+# Date: 24 March 2024
+# Prerequisites: None
+
+library(janitor)
+library(dplyr)
+library(arrow)
+
+raw_data <- clean_names(raw_data) 
+raw_data <- raw_data |> 
+  select(maternal_age, household_income, edinburgh_postnatal_depression_scale, promis_anxiety, delivery_date_converted_to_month_and_year, birth_length, birth_weight, 	
+         delivery_mode, language, threaten_life, threaten_baby_danger, threaten_baby_harm)
+
+raw_data_complete <- raw_data[complete.cases(raw_data), ]
+
+raw_data_complete <- raw_data_complete |> filter(household_income != "") |>
+  mutate(month = substr(delivery_date_converted_to_month_and_year, 1, 3),
+         year = substr(delivery_date_converted_to_month_and_year, 4, 7),
+         age_group = case_when(
+           maternal_age < 27 ~ 'Under 27',
+           maternal_age >= 27 & maternal_age < 35 ~ '27 to 34',
+           maternal_age >= 35 & maternal_age < 40 ~ '35 to 39',
+           maternal_age >= 40 ~ '40 and above'),
+         income = case_when(
+           household_income == "$200,000+" ~ "$200,000+",
+           household_income == "$100,000 -$124,999" ~ "$100,000 - $124,999",
+           household_income == "$40,000-$69,999" ~ "$40,000 - $69,999",
+           household_income == "$70,000-$99,999" ~ "$70,000 - $99,999",
+           household_income == "$20,000- $39,999" ~ "$20,000 - $39,999",
+           household_income == "$125,000- $149,999" ~ "$125,000 - $149,999",
+           household_income == "$150,000 - $174,999" ~ "$150,000 - $174,999",
+           household_income == "$175,000- $199,999" ~ "$175,000 - $199,999",
+           household_income == "Less than $20, 000" ~ "Less than $20,000"),
+         EPDS = edinburgh_postnatal_depression_scale/30,
+         PROMIS = promis_anxiety/35,
+         threaten_life = threaten_life/100,
+         threaten_baby_danger = threaten_baby_danger/100,
+         threaten_baby_harm = threaten_baby_harm/100)
+
[…]

### 2024-03-24 scripts/03-test.R

@@ -1,0 +1,8 @@
+library(arrow)
+library(readr)
+library(here)
+
+clean_data <- read_csv(
+  file = here("inputs/data/cleaned_mental_health.csv"),
+  show_col_types = FALSE
+)

### 2024-03-24 scripts/99-model.R

@@ -1,0 +1,48 @@
+#### Preamble ####
+# Purpose: Create models to estimate the level of depression and anxiety using stan_glm()
+# Author: Boxuan Yi
+# Email: boxuan.yi@mail.utoronto.ca
+# Date: 24 March 2024
+# Prerequisites: Have predictors
+
+library(rstanarm)
+library(here)
+
+clean_data <- read_csv(
+  file = here("inputs/data/cleaned_mental_health.csv"),
+  show_col_types = FALSE
+)
+
+# Depression Model
+depression_model <-
+  stan_glm(
+    EPDS ~ age_group + household_income + language + month,
+    data = clean_data,
+    family = binomial(link = "logit"),
+    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_intercept = 
+      normal(location = 0, scale = 2.5, autoscale = TRUE),
+    seed = 777
+  )
+
+saveRDS(
+  depression_model,
+  file = "model/depression.rds"
+)
+
+# Anxiety Model
+anxiety_model <-
+  stan_glm(
+    PROMIS ~ age_group + household_income + language + month,
+    data = clean_data,
+    family = binomial(link = "logit"),
+    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_intercept = 
+      normal(location = 0, scale = 2.5, autoscale = TRUE),
+    seed = 777
+  )
+
+saveRDS(
+  anxiety_model,
+  file = "model/anxiety.rds"
+)

### 2024-03-24 outputs/paper.qmd
render paper

@@
 ---
-title: "Untitled"
-author: "Boxuan Yi"
+title: ""
 format: pdf
+date: "24 March 2024"
+date-format: "D MMMM YYYY"
+number-sections: true
+thanks: "Code and data in this analysis is available at: https://github.com/Elaineyi1/Prenatal_Mental_Health"
+bibliography: references.bib
+toc: true
 ---
@@
 ```{r}
+#| include: false
 library(arrow)
@@
 ```{r}
+#| include: false
 # clean_data <- read_parquet(file = "inputs/data/cleaned_mental_health.parquet")
@@
     file = here("inputs/data/cleaned_mental_health.csv"),
-    show_col_types = FALSE
-  )
+    show_col_types = FALSE)
 ```
@@
 ```{r}
+#| include: false
 age_levels <- c("Under 27", "27 to 34", "35 to 39", "40 and above")
@@
 ```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-income_age
+#| fig-cap: The Distribution of Age Groups and Household Income for Pregnant Participants
 # Pastel1
@@
 ```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: tbl-income
+#| tbl-cap: Levels of Depression and Anxiety Among Different Income Groups
 clean_data |> group_by(household_income) |> 
@@
 ```{r}
-clean_data |> group_by(age_group) |> 
+#| message: false
+#| echo: false
+#| warning: false
+#| label: tbl-age
+#| tbl-cap: Levels of Depression and Anxiety Among Different Age Groups
+clean_data |> group_by(age_group, language) |> 
   summarise(average_depression = mean(EPDS), average_anxiety = mean(PROMIS)) |> 
   rename("Age Group" = age_group, 
+         "Survey Language" = language,
          "Average Level of Depression" = average_depression, 
@@
 ```{r}
+#| label: fig-month
+#| fig-cap: Distribution of Levels of Depression and Anxiety by Month
+#| echo: false
+#| warning: false
+#| message: false
+#| layout-ncol: 2
+#| fig-subcap: ["Distribution of Edinburgh Postnatal Depression Scale Scores by Month", "Distribution of Promis Anxiety Scores by Months"]
+clean_data |>
+  ggplot(aes(x = month, y = EPDS)) +
+  geom_boxplot() +
+  theme_classic()
+
+clean_data |>
+  ggplot(aes(x = month, y = PROMIS)) +
+  geom_boxplot() +
+  theme_classic()
+```
+
+# Model
+```{r}
+#| include: false
+depression_model <- readRDS(file = here::here("model/depression.rds"))
+```
+
+```{r}
+#| include: false
+anxiety_model <- readRDS(file = here::here("model/anxiety.rds"))
+```
+
+# Results
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: tbl-depression_model
+#| tbl-cap: Estimating Prenatal Depression given Age, Household Income, Month of Delivery and Choice of Language
+modelsummary(
+  list(
+    "Level of Depression" = depression_model))
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: tbl-anxiety_model
+#| tbl-cap: Estimating Prenatal Anxiety given Age Group, Household Income, Month of Delivery and Choice of Language.
+modelsummary(
+  list(
+    "Level of Anxiety" = anxiety_model))
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-depression_interval
+#| fig-cap: The Coefficients and Confidence Interval in the Depression Model
+modelplot(depression_model, conf_level = 0.95) +
+  labs(x = "95% Level of Depression Interval")
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-anxiety_interval
+#| fig-cap: The Coefficients and Confidence Interval in the Anxiety Model
+modelplot(anxiety_model, conf_level = 0.95) +
+  labs(x = "95% Level of Anxiety Interval")
+```
+
+```{r}
+#| label: fig-danger
+#| fig-cap: Association between Perceived Level of Danger and Prenatal Mental Health
+#| echo: false
+#| warning: false
+#| message: false
+#| layout-ncol: 2
+#| fig-subcap: ["Relationship between Perceived Danger to Self and Depression", "Relationship between Perceived Danger to the Unborn and Depression", "Relationship between Perceived Danger to Self and Anxiety", "Relationship between Perceived Danger to the Unborn and Anxiety"]
 clean_data |>
@@
 
-```{r}
-clean_data |>
-  ggplot(aes(x = month, y = EPDS)) +
-  geom_boxplot() +
-  theme_classic()
-
-clean_data |>
-  ggplot(aes(x = month, y = PROMIS)) +
-  geom_boxplot() +
-  theme_classic()
-```
-
-# Model
-```{r}
-depression_model <- readRDS(file = here::here("model/depression.rds"))
-```
-
-```{r}
-anxiety_model <- readRDS(file = here::here("model/anxiety.rds"))
-```
-
-# Results
-```{r}
-modelsummary(
-  list(
-    "Level of Depression" = depression_model))
-```
-
-```{r}
-modelsummary(
-  list(
-    "Level of Anxiety" = anxiety_model))
-```
-
-```{r}
-modelplot(depression_model, conf_level = 0.95) +
-  labs(x = "95% Level of Depression Interval")
-```
-
-```{r}
-modelplot(anxiety_model, conf_level = 0.95) +
-  labs(x = "95% Level of Anxiety Interval")
-```
-
 # Discussions
 ```{r}
+#| label: fig-weight_delivery
+#| fig-cap: Distribution of Birth Weight and Prenatal Mental Health by Delivery Mode
+#| echo: false
+#| warning: false
+#| message: false
+#| layout-ncol: 2
+#| fig-subcap: ["Distribution of Birth Weight and Depression by Delivery Mode", "Distribution of Birth Weight and Anxiety by Delivery Mode"]
 clean_data |>
@@
 ```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-danger_baby
+#| fig-cap: Distribution of the Preceived Danger to Baby and Preceived Potential Harm to Baby
 clean_data |>
@@
 ```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-language_income
+#| fig-cap: Income Group Distribution Based on the Choice of Survey Language
 clean_data |>
@@
   theme(legend.position = "bottom")
-
 ```

### 2024-03-25 outputs/paper.qmd
some improvements on figures

@@
 ---
-title: ""
-format: pdf
+title: "Analysis of Prenatal Depression and Anxiety during Covid-19"
+author: 
+  - Boxuan Yi
+format: 
+  pdf:
+    documentclass: article
+    geometry: margin = 1in
+    abstract: ""
 date: "24 March 2024"
@@
 ---
-
+# Introduction
 ```{r}
@@
 ```
+
+# Data
 
@@
   scale_fill_brewer(palette = "Set2") +
-  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
+  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
   labs(x = "Household Income in 2019", y = "Number of Observations", fill = "Age Group") 
 ```
+
 ```{r}
@@
   ggplot(aes(x = month, y = EPDS)) +
-  geom_boxplot() +
-  theme_classic()
+  geom_boxplot() + ylim(0,1)
 
@@
   ggplot(aes(x = month, y = PROMIS)) +
-  geom_boxplot() +
-  theme_classic()
+  geom_boxplot() + ylim(0,1)
 ```

### 2024-03-25 outputs/paper.qmd
finalize the tables, graphs and models

@@
 ---
-# Introduction
+# Introduction {#sec-introduction}
 ```{r}
@@
 
-# Data
+# Data {#sec-data}
 
@@
                    "$150,000 - $174,999", "$175,000 - $199,999", "$200,000+")
+education_levels <- c("Less than high school diploma", "High school diploma", "College/trade school", 
+                      "Undergraduate degree", "Masters degree", "Doctoral Degree")
 month_levels <- c("Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
@@
 clean_data$household_income <- factor(clean_data$household_income, levels = income_levels)
+clean_data$maternal_education <- factor(clean_data$maternal_education, levels = education_levels)
 clean_data$month <- factor(clean_data$month, levels = month_levels)
@@
 #| warning: false
-#| label: tbl-income
-#| tbl-cap: Levels of Depression and Anxiety Among Different Income Groups
-clean_data |> group_by(household_income) |> 
-  summarise(average_depression = mean(EPDS), average_anxiety = mean(PROMIS)) |> 
-  rename("Household Income in 2019" = household_income, 
-         "Average Level of Depression" = average_depression, 
-         "Average Level of Anxiety" = average_anxiety) |>
-  kable()
+#| label: fig-education_age
+#| fig-cap: The Distribution of Age Groups and Level of Education for Pregnant Participants
+clean_data |>
+  ggplot(mapping = aes(x = maternal_education, fill = age_group)) +
+  geom_bar(position = "dodge2") + 
+  theme(legend.position = "bottom") +
+  scale_fill_brewer(palette = "Set2") +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
+  labs(x = "Individual's Education Level", y = "Number of Observations", fill = "Age Group")
 ```
@@
 #| warning: false
-#| label: tbl-age
+#| label: tbl-age_language
 #| tbl-cap: Levels of Depression and Anxiety Among Different Age Groups
@@
   kable()
+```
+
+```{r}
+#| label: fig-average_depression
+#| fig-cap: Distribution of Levels of Depression by Education and Household Income
+#| echo: false
+#| warning: false
+#| message: false
+#| layout-ncol: 2
+#| fig-subcap: ["Distribution of Levels of Depression by Education", "Distribution of Levels of Depression by Household Income"]
+clean_data %>%
+  group_by(maternal_education) %>%
+  summarise(average_depression = mean(EPDS)) %>%
+  ggplot(aes(x = maternal_education, y = average_depression)) +
+  geom_bar(stat = "identity", fill = "cornflowerblue") + geom_text(aes(label = round(average_depression, 3)), vjust = -0.5) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.5) +
+  labs(x="Education", y="Average Level of Depression")
+
+clean_data %>%
+  group_by(household_income) %>%
+  summarise(average_depression = mean(EPDS)) %>%
+  ggplot(aes(x = household_income, y = average_depression)) + 
+  geom_text(aes(label = round(average_depression, 3)), vjust = -0.5) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.5)+
+  geom_bar(stat = "identity", fill = "cornflowerblue") + labs(x="Household Income in 2019", y="Average Level of Depression")
+```
+
+```{r}
+#| label: fig-average_anxiety
+#| fig-cap: Distribution of Levels of Anxiety by Education and Household Income
+#| echo: false
+#| warning: false
+#| message: false
+#| layout-ncol: 2
+#| fig-subcap: ["Distribution of Levels of Anxiety by Education", "Distribution of Levels of Anxiety by Household Income"]
+clean_data %>%
+  group_by(maternal_education) %>%
+  summarise(average_anxiety = mean(PROMIS)) %>%
+  ggplot(aes(x = maternal_education, y = average_anxiety)) +
+  geom_bar(stat = "identity", fill = "salmon2") + geom_text(aes(label = round(average_anxiety, 3)), vjust = -0.5) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.7)+
+  labs(x="Education", y="Average Level of Anxiety")
+
+clean_data %>%
+  group_by(household_income) %>%
+  summarise(average_anxiety = mean(PROMIS)) %>%
+  ggplot(aes(x = household_income, y = average_anxiety)) + 
+  geom_bar(stat = "identity", fill = "salmon2") +
+  geom_text(aes(label = round(average_anxiety, 3)), vjust = -0.5) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.7) +
+  labs(x="Household Income in 2019", y="Average Level of Anxiety")
 ```
@@
 
-# Model
+# Model {#sec-model}
 ```{r}
@@
 
-# Results
+# Results {#sec-results}
 ```{r}
@@
 #| fig-cap: The Coefficients and Confidence Interval in the Depression Model
-modelplot(depression_model, conf_level = 0.95) +
+modelplot(depression_model, conf_level = 0.95, color = 'royalblue3') +
   labs(x = "95% Level of Depression Interval")
@@
 #| fig-cap: The Coefficients and Confidence Interval in the Anxiety Model
-modelplot(anxiety_model, conf_level = 0.95) +
+modelplot(anxiety_model, conf_level = 0.95, color = 'indianred3') +
   labs(x = "95% Level of Anxiety Interval")
+```
+
+# Discussions {#sec-discussions}
+
+```{r}
+#| label: fig-weight_delivery
+#| fig-cap: Distribution of Birth Weight and Length and Prenatal Mental Health by Delivery Mode
+#| echo: false
+#| warning: false
+#| message: false
+#| layout-ncol: 2
+#| fig-subcap: ["Distribution of Birth Weight and Depression by Delivery Mode", "Distribution of Birth Weight and Anxiety by Delivery Mode", "Distribution of Birth Length and Depression by Delivery Mode", "Distribution of Birth Length and Anxiety by Delivery Mode"]
+clean_data |>
+  ggplot(aes(x = EPDS, y = birth_weight, color = delivery_mode)) +
+  geom_smooth(method = "lm")+
+  theme_minimal() + labs(x = "The Level of Prenatal Depression", y = "Birth Weight (grams)")
+
+clean_data |>
+  ggplot(aes(x = PROMIS, y = birth_weight, color = delivery_mode)) +
+  geom_smooth(method = "lm")+
+  theme_minimal() +
+  labs(x = "The Level of Prenatal Anxiety", y = "Birth Weight (grams)")
+
+clean_data |>
+  ggplot(aes(x = EPDS, y = birth_length, color = delivery_mode)) +
+  geom_smooth(method = "lm")+
+  theme_minimal() + labs(x = "The Level of Prenatal Depression", y = "Birth Length (centimeters)")
+
+clean_data |>
+  ggplot(aes(x = PROMIS, y = birth_length, color = delivery_mode)) +
+  geom_smooth(method = "lm")+
+  theme_minimal() +
+  labs(x = "The Level of Prenatal Anxiety", y = "Birth Length (centimeters)")
+```
+
+# Appendix {#sec-appendix}
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-language_income_education
+#| fig-cap: Income and Education Groups Distribution Based on the Choice of Survey Language
+#| layout-ncol: 2
+#| fig-subcap: ["Distribution of Income by Language", "Distribution of Education by Language"]
+clean_data |>
+  ggplot(aes(x = household_income, fill = language)) +
+  stat_count(position = "dodge") +
+  theme_minimal() +
+  coord_flip() +
+  scale_fill_brewer(palette = "Set1") + theme(legend.position = "bottom")
+
+clean_data |>
+  ggplot(aes(x = maternal_education, fill = language)) +
+  stat_count(position = "dodge") +
+  theme_minimal() +
+  coord_flip() +
+  scale_fill_brewer(palette = "Set1") + theme(legend.position = "bottom")
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-danger_baby
+#| fig-cap: Distribution of the Preceived Danger to Baby and Preceived Potential Harm to Baby
+clean_data |>
+  ggplot(aes(x = threaten_baby_danger, y = threaten_baby_harm)) +
+  geom_jitter(color='darkgray', alpha=0.4) +
+  geom_smooth() + theme_minimal()
 ```
@@
   geom_smooth(method = "lm") + 
+  geom_hline(yintercept = 0.3, linetype = "dashed", color = "black") +
   theme_classic() + 
-  labs(x="Perceived Level of Danger", y = "Level of Prenatal Depression") + ylim(0,1)
+  labs(x="Perceived Level of Danger", y = "Level of Prenatal Depression")+ ylim(0,1)
 
@@
   geom_smooth(method = "lm") + 
+  geom_hline(yintercept = 0.3, linetype = "dashed", color = "black") +
   theme_classic() +
@@
   geom_smooth(method = "lm") +
-  theme_classic() +
-  labs(x="Perceived Level of Danger", y = "Level of Prenatal Anxiety") + ylim(0,1)
+  theme_classic() + geom_hline(yintercept = c(0.4, 0.6), linetype = "dashed", color = "black")+
+  labs(x="Perceived Level of Danger", y = "Level of Prenatal Anxiety")+ ylim(0,1)
 
@@
   geom_smooth(method = "lm") + 
+  geom_hline(yintercept = c(0.4, 0.6), linetype = "dashed", color = "black") +
   theme_classic() +
@@
 ```
-
-# Discussions
-```{r}
-#| label: fig-weight_delivery
-#| fig-cap: Distribution of Birth Weight and Prenatal Mental Health by Delivery Mode
-#| echo: false
-#| warning: false
-#| message: false
-#| layout-ncol: 2
-#| fig-subcap: ["Distribution of Birth Weight and Depression by Delivery Mode", "Distribution of Birth Weight and Anxiety by Delivery Mode"]
-clean_data |>
-  ggplot(aes(x = EPDS, y = birth_weight, color = delivery_mode)) +
-  geom_smooth(method = "lm")+
-  theme_minimal() + labs(x = "The Level of Prenatal Depression", y = "Birth Weight (grams)")
-
-clean_data |>
-  ggplot(aes(x = PROMIS, y = birth_weight, color = delivery_mode)) +
-  geom_smooth(method = "lm")+
-  theme_minimal() +
-  labs(x = "The Level of Prenatal Anxiety", y = "Birth Weight (grams)")
-```
-
-# Appendix
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-danger_baby
-#| fig-cap: Distribution of the Preceived Danger to Baby and Preceived Potential Harm to Baby
-clean_data |>
-  ggplot(aes(x = threaten_baby_danger, y = threaten_baby_harm)) +
-  geom_jitter(color='darkgray', alpha=0.4) +
-  geom_smooth() +
-  theme_classic() 
-```
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-language_income
-#| fig-cap: Income Group Distribution Based on the Choice of Survey Language
-clean_data |>
-  ggplot(aes(x = household_income, fill = language)) +
-  stat_count(position = "dodge") +
-  theme_minimal() +
-  coord_flip() +
-  scale_fill_brewer(palette = "Set1") +
-  theme(legend.position = "bottom")
-```

### 2024-03-25 scripts/02-clean.R

@@ -11,12 +11,14 @@ library(arrow)
 
 raw_data <- clean_names(raw_data) 
 raw_data <- raw_data |> 
-  select(maternal_age, household_income, edinburgh_postnatal_depression_scale, promis_anxiety, delivery_date_converted_to_month_and_year, birth_length, birth_weight, 	
+  select(maternal_age, household_income, edinburgh_postnatal_depression_scale, promis_anxiety, 
+         delivery_date_converted_to_month_and_year, birth_length, birth_weight, maternal_education, 
          delivery_mode, language, threaten_life, threaten_baby_danger, threaten_baby_harm)
 
 raw_data_complete <- raw_data[complete.cases(raw_data), ]
 
-raw_data_complete <- raw_data_complete |> filter(household_income != "") |>
+raw_data_complete <- raw_data_complete |> filter(household_income != "") |> 
+  filter(maternal_education != "") |>
   mutate(month = substr(delivery_date_converted_to_month_and_year, 1, 3),
          year = substr(delivery_date_converted_to_month_and_year, 4, 7),
          age_group = case_when(
@@ -41,7 +43,9 @@ raw_data_complete <- raw_data_complete |> filter(household_income != "") |>
          threaten_baby_harm = threaten_baby_harm/100)
 
 clean_data <- raw_data_complete |>  filter(year == '2020') |>
-  select(age_group, maternal_age, income, birth_weight, delivery_mode, language, threaten_life, threaten_baby_danger, threaten_baby_harm, month, EPDS, PROMIS) |>
+  select(age_group, maternal_age, income, maternal_education, birth_weight, birth_length,
+         delivery_mode, language, threaten_life, threaten_baby_danger, threaten_baby_harm, 
+         month, EPDS, PROMIS) |>
   rename(household_income = income)
 
 write.csv(

### 2024-03-25 scripts/99-model.R

@@ -16,7 +16,7 @@ clean_data <- read_csv(
 # Depression Model
 depression_model <-
   stan_glm(
-    EPDS ~ age_group + household_income + language + month,
+    EPDS ~ age_group + household_income + language + maternal_education,
     data = clean_data,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
@@ -33,7 +33,7 @@ saveRDS(
 # Anxiety Model
 anxiety_model <-
   stan_glm(
-    PROMIS ~ age_group + household_income + language + month,
+    PROMIS ~ age_group + household_income + language + maternal_education,
     data = clean_data,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
@@ -46,3 +46,5 @@ saveRDS(
   anxiety_model,
   file = "model/anxiety.rds"
 )
+
+

### 2024-03-25 outputs/paper.qmd
correct some captions

@@
 #| label: tbl-depression_model
-#| tbl-cap: Estimating Prenatal Depression given Age, Household Income, Month of Delivery and Choice of Language
+#| tbl-cap: Estimating Prenatal Depression given Age, Household Income, Education Level and Choice of Language
 modelsummary(
@@
 #| label: tbl-anxiety_model
-#| tbl-cap: Estimating Prenatal Anxiety given Age Group, Household Income, Month of Delivery and Choice of Language.
+#| tbl-cap: Estimating Prenatal Anxiety given Age Group, Household Income, Education Level and Choice of Language.
 modelsummary(
@@
 modelplot(depression_model, conf_level = 0.95, color = 'royalblue3') +
-  labs(x = "95% Level of Depression Interval")
+  labs(x = "Estimating Depression and 95% Confidence Interval")
 ```
@@
 modelplot(anxiety_model, conf_level = 0.95, color = 'indianred3') +
-  labs(x = "95% Level of Anxiety Interval")
+  labs(x = "Estimating Anxiety and 95% Confidence Interval")
 ```
@@
 ```{r}
-#| label: fig-weight_delivery
+#| label: fig-weight_length_delivery
 #| fig-cap: Distribution of Birth Weight and Length and Prenatal Mental Health by Delivery Mode

### 2024-03-25 scripts/02-clean.R
filter the dataset

@@ -42,7 +42,7 @@ raw_data_complete <- raw_data_complete |> filter(household_income != "") |>
          threaten_baby_danger = threaten_baby_danger/100,
          threaten_baby_harm = threaten_baby_harm/100)
 
-clean_data <- raw_data_complete |>  filter(year == '2020') |>
+clean_data <- raw_data_complete |>  filter(year == '2020') |> filter(month != 'Apr') |>
   select(age_group, maternal_age, income, maternal_education, birth_weight, birth_length,
          delivery_mode, language, threaten_life, threaten_baby_danger, threaten_baby_harm, 
          month, EPDS, PROMIS) |>

### 2024-03-25 scripts/99-model.R

@@ -13,6 +13,20 @@ clean_data <- read_csv(
   show_col_types = FALSE
 )
 
+# Setting levels
+age_levels <- c("Under 27", "27 to 34", "35 to 39", "40 and above")
+income_levels <- c("Less than $20,000", "$20,000 - $39,999", "$40,000 - $69,999", 
+                   "$70,000 - $99,999", "$100,000 - $124,999", "$125,000 - $149,999",
+                   "$150,000 - $174,999", "$175,000 - $199,999", "$200,000+")
+education_levels <- c("Less than high school diploma", "High school diploma", "College/trade school", 
+                      "Undergraduate degree", "Masters degree", "Doctoral Degree")
+month_levels <- c("Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
+clean_data$age_group <- factor(clean_data$age_group, levels = age_levels)
+clean_data$household_income <- factor(clean_data$household_income, levels = income_levels)
+clean_data$maternal_education <- factor(clean_data$maternal_education, levels = education_levels)
+clean_data$month <- factor(clean_data$month, levels = month_levels)
+
+
 # Depression Model
 depression_model <-
   stan_glm(

### 2024-03-27 outputs/paper.qmd
modify the model

@@
 ```{r}
+#| label: fig-month
+#| fig-cap: Distribution of Levels of Depression and Anxiety by Month
+#| echo: false
+#| warning: false
+#| message: false
+#| layout-ncol: 2
+#| fig-subcap: ["Distribution of Edinburgh Postnatal Depression Scale Scores by Month", "Distribution of Promis Anxiety Scores by Months"]
+clean_data |>
+  ggplot(aes(x = month, y = EPDS)) +
+  geom_boxplot() + ylim(0,1)
+
+clean_data |>
+  ggplot(aes(x = month, y = PROMIS)) +
+  geom_boxplot() + ylim(0,1)
+```
+
+```{r}
 #| message: false
@@
 ```{r}
-#| label: fig-average_depression
-#| fig-cap: Distribution of Levels of Depression by Education and Household Income
+#| label: fig-average_depression_and_anxiety
+#| fig-cap: Distribution of Levels of Depression and Anxiety by Education and Household Income
 #| echo: false
@@
 #| layout-ncol: 2
-#| fig-subcap: ["Distribution of Levels of Depression by Education", "Distribution of Levels of Depression by Household Income"]
+#| fig-subcap: ["Distribution of Levels of Depression by Education", "Distribution of Levels of Depression by Household Income", "Distribution of Levels of Anxiety by Education", "Distribution of Levels of Anxiety by Household Income"]
 clean_data %>%
@@
   geom_bar(stat = "identity", fill = "cornflowerblue") + geom_text(aes(label = round(average_depression, 3)), vjust = -0.5) +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.5) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.7) +
   labs(x="Education", y="Average Level of Depression")
@@
   geom_text(aes(label = round(average_depression, 3)), vjust = -0.5) +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.5)+
+  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.7)+
   geom_bar(stat = "identity", fill = "cornflowerblue") + labs(x="Household Income in 2019", y="Average Level of Depression")
-```
 
-```{r}
-#| label: fig-average_anxiety
-#| fig-cap: Distribution of Levels of Anxiety by Education and Household Income
-#| echo: false
-#| warning: false
-#| message: false
-#| layout-ncol: 2
-#| fig-subcap: ["Distribution of Levels of Anxiety by Education", "Distribution of Levels of Anxiety by Household Income"]
 clean_data %>%
@@
 ```{r}
-#| label: fig-month
-#| fig-cap: Distribution of Levels of Depression and Anxiety by Month
+#| label: fig-danger
+#| fig-cap: Relationship between Perceived Level of Danger and Prenatal Mental Health
 #| echo: false
@@
 #| layout-ncol: 2
-#| fig-subcap: ["Distribution of Edinburgh Postnatal Depression Scale Scores by Month", "Distribution of Promis Anxiety Scores by Months"]
+#| fig-subcap: ["Perceived Danger to Self and Depression", "Perceived Danger to the Unborn and Depression", "Perceived Danger to Self and Anxiety", "Perceived Danger to the Unborn and Anxiety"]
 clean_data |>
-  ggplot(aes(x = month, y = EPDS)) +
-  geom_boxplot() + ylim(0,1)
+  ggplot(aes(x = threaten_life, y = EPDS, color = language)) +
+  geom_smooth(method = "lm") + 
+  geom_hline(yintercept = 0.3, linetype = "dashed", color = "black") +
+  theme_classic() + 
+  geom_text(aes(x = 1, y = 0.3, label = "0.3"), color = "black", vjust = -0.5, hjust = 1) +
+  labs(x="Perceived Level of Danger", y = "Level of Prenatal Depression")+ ylim(0,1)
 
 clean_data |>
-  ggplot(aes(x = month, y = PROMIS)) +
-  geom_boxplot() + ylim(0,1)
+  ggplot(aes(x = threaten_baby_danger, y = EPDS, color = language)) +
+  geom_smooth(method = "lm") + 
+  geom_hline(yintercept = 0.3, linetype = "dashed", color = "black") +
+  theme_classic() + 
+  geom_text(aes(x = 1, y = 0.3, label = "0.3"), color = "black", vjust = -0.5, hjust = 1) +
+  labs(x="Perceived Danger to Unborn Baby's Life", y = "Level of Prenatal Depression") + ylim(0,1)
+
+clean_data |>
+  ggplot(aes(x = threaten_life, y = PROMIS, color = language)) +
+  geom_smooth(method = "lm") +
+  theme_classic() + geom_hline(yintercept = 0.5, linetype = "dashed", color = "black")+
+  geom_text(aes(x = 1, y = 0.5, label = "0.5"), color = "black", vjust = -0.5, hjust = 1) +
+  labs(x="Perceived Level of Danger", y = "Level of Prenatal Anxiety")+ ylim(0,1)
+
+clean_data |>
+  ggplot(aes(x = threaten_baby_danger, y = PROMIS, color = language)) +
+  geom_smooth(method = "lm") + 
+  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black") +
+  theme_classic() + 
+  geom_text(aes(x = 1, y = 0.5, label = "0.5"), color = "black", vjust = -0.5, hjust = 1) +
+  labs(x="Perceived Danger to Unborn Baby's Life", y = "Level of Prenatal Anxiety") + ylim(0,1) 
 ```
@@
   list(
-    "Level of Depression" = depression_model))
+    "Depression" = depression_model))
+```
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-depression_interval
+#| fig-cap: The Coefficients and Confidence Interval in the Depression Model
+modelplot(depression_model, conf_level = 0.95, color = 'royalblue3') +
+  labs(x = "Depression and 95% Confidence Interval")
 ```
@@
   list(
-    "Level of Anxiety" = anxiety_model))
-```
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-depression_interval
-#| fig-cap: The Coefficients and Confidence Interval in the Depression Model
-modelplot(depression_model, conf_level = 0.95, color = 'royalblue3') +
-  labs(x = "Estimating Depression and 95% Confidence Interval")
+    "Anxiety" = anxiety_model))
 ```
@@
 modelplot(anxiety_model, conf_level = 0.95, color = 'indianred3') +
-  labs(x = "Estimating Anxiety and 95% Confidence Interval")
+  labs(x = "Anxiety and 95% Confidence Interval")
 ```
@@
 # Discussions {#sec-discussions}
-
 ```{r}
@@
 #| layout-ncol: 2
-#| fig-subcap: ["Distribution of Birth Weight and Depression by Delivery Mode", "Distribution of Birth Weight and Anxiety by Delivery Mode", "Distribution of Birth Length and Depression by Delivery Mode", "Distribution of Birth Length and Anxiety by Delivery Mode"]
+#| fig-subcap: ["Birth Weight and Depression by Delivery Mode", "Birth Weight and Anxiety by Delivery Mode", "Birth Length and Depression by Delivery Mode", "Birth Length and Anxiety by Delivery Mode"]
 clean_data |>
@@
 ```
-
-```{r}
-#| label: fig-danger
-#| fig-cap: Association between Perceived Level of Danger and Prenatal Mental Health
-#| echo: false
-#| warning: false
-#| message: false
-#| layout-ncol: 2
-#| fig-subcap: ["Relationship between Perceived Danger to Self and Depression", "Relationship between Perceived Danger to the Unborn and Depression", "Relationship between Perceived Danger to Self and Anxiety", "Relationship between Perceived Danger to the Unborn and Anxiety"]
-clean_data |>
-  ggplot(aes(x = threaten_life, y = EPDS, color = language)) +
-  geom_smooth(method = "lm") + 
-  geom_hline(yintercept = 0.3, linetype = "dashed", color = "black") +
-  theme_classic() + 
-  labs(x="Perceived Level of Danger", y = "Level of Prenatal Depression")+ ylim(0,1)
-
-clean_data |>
-  ggplot(aes(x = threaten_baby_danger, y = EPDS, color = language)) +
-  geom_smooth(method = "lm") + 
-  geom_hline(yintercept = 0.3, linetype = "dashed", color = "black") +
-  theme_classic() +
-  labs(x="Perceived Danger to Unborn Baby's Life", y = "Level of Prenatal Depression") + ylim(0,1)
-
-clean_data |>
-  ggplot(aes(x = threaten_life, y = PROMIS, color = language)) +
-  geom_smooth(method = "lm") +
-  theme_classic() + geom_hline(yintercept = c(0.4, 0.6), linetype = "dashed", color = "black")+
-  labs(x="Perceived Level of Danger", y = "Level of Prenatal Anxiety")+ ylim(0,1)
-
-clean_data |>
-  ggplot(aes(x = threaten_baby_danger, y = PROMIS, color = language)) +
-  geom_smooth(method = "lm") + 
-  geom_hline(yintercept = c(0.4, 0.6), linetype = "dashed", color = "black") +
-  theme_classic() +
-  labs(x="Perceived Danger to Unborn Baby's Life", y = "Level of Prenatal Anxiety") + ylim(0,1) 
-```

### 2024-03-27 scripts/99-model.R

@@ -30,7 +30,7 @@ clean_data$month <- factor(clean_data$month, levels = month_levels)
 # Depression Model
 depression_model <-
   stan_glm(
-    EPDS ~ age_group + household_income + language + maternal_education,
+    EPDS ~ age_group + household_income + language + maternal_education + threaten_life + threaten_baby_danger,
     data = clean_data,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
@@ -47,7 +47,7 @@ saveRDS(
 # Anxiety Model
 anxiety_model <-
   stan_glm(
-    PROMIS ~ age_group + household_income + language + maternal_education,
+    PROMIS ~ age_group + household_income + language + maternal_education + threaten_life + threaten_baby_danger,
     data = clean_data,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
@@ -60,5 +60,4 @@ saveRDS(
   anxiety_model,
   file = "model/anxiety.rds"
 )
-
 

### 2024-03-27 scripts/02-clean.R
create tests, modify a variable

@@ -37,7 +37,7 @@ raw_data_complete <- raw_data_complete |> filter(household_income != "") |>
            household_income == "$175,000- $199,999" ~ "$175,000 - $199,999",
            household_income == "Less than $20, 000" ~ "Less than $20,000"),
          EPDS = edinburgh_postnatal_depression_scale/30,
-         PROMIS = promis_anxiety/35,
+         PROMIS = (promis_anxiety-7)/28,
          threaten_life = threaten_life/100,
          threaten_baby_danger = threaten_baby_danger/100,
          threaten_baby_harm = threaten_baby_harm/100)

### 2024-03-27 scripts/03-test.R

@@ -6,3 +6,33 @@ clean_data <- read_csv(
   file = here("inputs/data/cleaned_mental_health.csv"),
   show_col_types = FALSE
 )
+
+# Maximum and minimum of scores
+max(clean_data$PROMIS) <= 1
+max(clean_data$EPDS) <= 1
+max(clean_data$threaten_life) <= 1
+max(clean_data$threaten_baby_danger) <= 1
+max(clean_data$threaten_baby_harm) <= 1
+
+min(clean_data$PROMIS) >= 0
+min(clean_data$EPDS) >= 0
+min(clean_data$threaten_life) >= 0
+min(clean_data$threaten_baby_danger) >= 0
+min(clean_data$threaten_baby_harm) >= 0
+
+# Groups and classes
+length(unique(clean_data$age_group)) <= 4
+length(unique(clean_data$household_income)) <= 9
+length(unique(clean_data$maternal_education)) <= 6
+is.numeric(clean_data$maternal_age)
+is.numeric(clean_data$birth_weight)
+is.numeric(clean_data$birth_length)
+is.numeric(clean_data$EPDS)
+is.numeric(clean_data$PROMIS)
+
+unique(clean_data$language) %in% c("English", "French")
+length(unique(clean_data$month)) <= 8
+length(unique(clean_data$delivery_mode)) == 2
+
+class(clean_data$language) == "character"
+class(clean_data$delivery_mode) == "character"

### 2024-03-27 outputs/paper.qmd
modify the anxiety model based the modification of the anxiety variable

@@
   geom_bar(stat = "identity", fill = "cornflowerblue") + geom_text(aes(label = round(average_depression, 3)), vjust = -0.5) +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.7) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.6) +
   labs(x="Education", y="Average Level of Depression")
@@
   geom_text(aes(label = round(average_depression, 3)), vjust = -0.5) +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.7)+
+  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.6)+
   geom_bar(stat = "identity", fill = "cornflowerblue") + labs(x="Household Income in 2019", y="Average Level of Depression")
@@
   geom_bar(stat = "identity", fill = "salmon2") + geom_text(aes(label = round(average_anxiety, 3)), vjust = -0.5) +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.7)+
+  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.6)+
   labs(x="Education", y="Average Level of Anxiety")
@@
   geom_text(aes(label = round(average_anxiety, 3)), vjust = -0.5) +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.7) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.6) +
   labs(x="Household Income in 2019", y="Average Level of Anxiety")
@@
   geom_smooth(method = "lm") +
-  theme_classic() + geom_hline(yintercept = 0.5, linetype = "dashed", color = "black")+
-  geom_text(aes(x = 1, y = 0.5, label = "0.5"), color = "black", vjust = -0.5, hjust = 1) +
+  theme_classic() + geom_hline(yintercept = 0.4, linetype = "dashed", color = "black")+
+  geom_text(aes(x = 1, y = 0.4, label = "0.4"), color = "black", vjust = -0.5, hjust = 1) +
   labs(x="Perceived Level of Danger", y = "Level of Prenatal Anxiety")+ ylim(0,1)
@@
   geom_smooth(method = "lm") + 
-  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black") +
+  geom_hline(yintercept = 0.4, linetype = "dashed", color = "black") +
   theme_classic() + 
-  geom_text(aes(x = 1, y = 0.5, label = "0.5"), color = "black", vjust = -0.5, hjust = 1) +
+  geom_text(aes(x = 1, y = 0.4, label = "0.4"), color = "black", vjust = -0.5, hjust = 1) +
   labs(x="Perceived Danger to Unborn Baby's Life", y = "Level of Prenatal Anxiety") + ylim(0,1) 

### 2024-03-27 outputs/paper.qmd
add data section

@@
 ---
-title: "Analysis of Prenatal Depression and Anxiety during Covid-19"
+title: "Exploring Depression and Anxiety: Objective Hardship Exposure and Psychological Distress Among Pregnant Individuals"
+subtitle: "Analysis of Prenatal Mental Health During the Covid-19 Pandemic"
 author: 
@@
     geometry: margin = 1in
-    abstract: ""
-date: "24 March 2024"
+    abstract: "The outbreak of the Covid-19 pandemic was a substantial stressor, especially for pregnant individuals.
+This paper uses Bayesian model and logistic regression to estimate levels of prenatal depression and anxiety during this hardship.
+The analysis reveals that the perceived threats to oneself and the unborn child are positively correlated to the mental distress, especially among younger pregnant individuals with lower levels of education and from lower-income families."
+date: "27 March 2024"
 date-format: "D MMMM YYYY"
@@
 ---
+\newpage
 # Introduction {#sec-introduction}
@@
 # Data {#sec-data}
+## Data Source and Measurement
+The prenatal mental health dataset from the Open Science Framework was designed to explore the relationship between exposure to objective hardship caused by the Covid -19 pandemic and psychological distress in pregnant individuals [@osfPregnancyDuring].
+The  study population is pregnant individuals aged 17 or older, with gestation periods of no more than 35 weeks, residing in Canada.
+Data were collected from May 2020 to December 2021, containing demographic information, mental health data and basic birth outcomes [@nihPrenatalMental].
+For this paper, I will only use data collected from May 2020 to December 2020 to analyze the prenatal mental health during the onset of the outbreak, a period highly marked by panic and hardship.
+
+Participants were recruited through advertisements via pregnancy organizations, care providers, social media, and paid ads on Facebook and Instagram, with the chance to win a $500 gift card.
+A portion of advertisements targeted geographic regions and sociodemographic groups with less representation to reduce underrepresentation [@nihPrenatalMental].
+
+Prenatal data, including age, 2019 household income in Canadian dollars, and education level, were collected as part of the national Pregnancy during the COVID-19 Pandemic (PdP) project using online questionnaires through REDCap.
+Birth data were also acquired using parents reports in REDCap [@nihPrenatalMental].
+Depression symptoms were self-assessed using the Edinburgh Postnatal Depression Scale (EPDS), consisting of 10 questions, each scored from 0 to 3, with possible scores ranging from 0 to 30.
+Anxiety symptoms were self-assessed by Patient-Reported Outcomes Measurement Information System (PROMIS), including 7 questions, each scored from 1 to 5, with possible total scores ranging from 7 to 35.
+Higher scores on both surveys indicate more severe depression or anxiety symptoms.
+The surveys were available in either English or French.
+
+Data are available for 10,772 participants.
+After excluding observations with missing values and setting the delivery month from May 2020 to December 2020, 3266 observations remain.
+Further details regarding data cleaning can be found in @sec-appendix.
+All participants in this study provided informed consent.
 
@@
 ```
-
+\newpage
 # Appendix {#sec-appendix}
@@
 ```
+\newpage
+# References

### 2024-03-29 outputs/paper.qmd
fix the model

@@
 ---
-title: "Exploring Depression and Anxiety: Objective Hardship Exposure and Psychological Distress Among Pregnant Individuals"
-subtitle: "Analysis of Prenatal Mental Health During the Covid-19 Pandemic"
+title: "Depression, Anxiety, and Objective Hardship: Examining Prenatal Mental Health During the Covid-19 Pandemic in Canada"
 author: 
@@
 A portion of advertisements targeted geographic regions and sociodemographic groups with less representation to reduce underrepresentation [@nihPrenatalMental].
-
 Prenatal data, including age, 2019 household income in Canadian dollars, and education level, were collected as part of the national Pregnancy during the COVID-19 Pandemic (PdP) project using online questionnaires through REDCap.
@@
 Higher scores on both surveys indicate more severe depression or anxiety symptoms.
+In this paper, the EPDS scores are divided by 30, and the PROMIS scores are adjusted by subtracting 7 from the original scores and then dividing by 35, and hence, making the level of depression and anxiety between 0 and 1 to standardize the scale.
 The surveys were available in either English or French.
@@
 ```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: tbl-age_language
-#| tbl-cap: Levels of Depression and Anxiety Among Different Age Groups
-clean_data |> group_by(age_group, language) |> 
-  summarise(average_depression = mean(EPDS), average_anxiety = mean(PROMIS)) |> 
-  rename("Age Group" = age_group, 
-         "Survey Language" = language,
-         "Average Level of Depression" = average_depression, 
-         "Average Level of Anxiety" = average_anxiety) |>
-  kable()
-```
-
-```{r}
-#| label: fig-average_depression_and_anxiety
-#| fig-cap: Distribution of Levels of Depression and Anxiety by Education and Household Income
+#| label: fig-age_language_dep_anx
+#| fig-cap: Levels of Depression and Anxiety Among Pregnant Individuals Grouped By Age and Language
 #| echo: false
@@
 #| layout-ncol: 2
-#| fig-subcap: ["Distribution of Levels of Depression by Education", "Distribution of Levels of Depression by Household Income", "Distribution of Levels of Anxiety by Education", "Distribution of Levels of Anxiety by Household Income"]
-clean_data %>%
-  group_by(maternal_education) %>%
-  summarise(average_depression = mean(EPDS)) %>%
-  ggplot(aes(x = maternal_education, y = average_depression)) +
-  geom_bar(stat = "identity", fill = "cornflowerblue") + geom_text(aes(label = round(average_depression, 3)), vjust = -0.5) +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.6) +
-  labs(x="Education", y="Average Level of Depression")
+#| fig-subcap: ["Age and Depression", "Language and Depression", "Age and Anxiety", "Language and Anxiety"]
+clean_data |> 
+  ggplot(aes(x=age_group,fill = factor(depression))) +
+  geom_bar(position = "fill") +
+  scale_fill_manual(values = c("0" = "lightcyan3", "1" = "cadetblue"), 
+                    labels = c("No or Mild Depression", "Moderate Depression or more"))+
+  coord_flip()+
+  labs(fill = "Depression", x = "Age Group", y = "Proportion") +
+  guides(fill = guide_legend(position = "bottom"))
 
-clean_data %>%
-  group_by(household_income) %>%
-  summarise(average_depression = mean(EPDS)) %>%
-  ggplot(aes(x = household_income, y = average_depression)) + 
-  geom_text(aes(label = round(average_depression, 3)), vjust = -0.5) +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.6)+
-  geom_bar(stat = "identity", fill = "cornflowerblue") + labs(x="Household Income in 2019", y="Average Level of Depression")
+clean_data |>
+  ggplot(aes(x = language, fill = factor(depression))) +
+  geom_bar(position = "fill") +
+  scale_fill_manual(values = c("0" = "lightcyan3", "1" = "cadetblue"), 
+                    labels = c("No or Mild Depression", "Moderate Depression or more")) +
+  coord_flip() +
+  labs(fill = "Depression", x = "Language of the Survey", y = "Proportion") +
+  guides(fill = guide_legend(position = "bottom"))
 
-clean_data %>%
-  group_by(maternal_education) %>%
-  summarise(average_anxiety = mean(PROMIS)) %>%
-  ggplot(aes(x = maternal_education, y = average_anxiety)) +
-  geom_bar(stat = "identity", fill = "salmon2") + geom_text(aes(label = round(average_anxiety, 3)), vjust = -0.5) +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.6)+
-  labs(x="Education", y="Average Level of Anxiety")
+clean_data |>
+  ggplot(aes(x = age_group, fill = factor(anxiety))) +
+  geom_bar(position = "fill") +
+  scale_fill_manual(values = c("0" = "#f4bfb6", "1" = "#f09286"), 
+                    labels = c("No or Mild Anxiety", "Moderately Anxiety or more")) +
+  coord_flip() +
+  labs(fill = "Anxiety", x = "Age Group", y = "Proportion")+
+  guides(fill = guide_legend(position = "bottom"))
 
-clean_data %>%
-  group_by(household_income) %>%
-  summarise(average_anxiety = mean(PROMIS)) %>%
-  ggplot(aes(x = household_income, y = average_anxiety)) + 
-  geom_bar(stat = "identity", fill = "salmon2") +
-  geom_text(aes(label = round(average_anxiety, 3)), vjust = -0.5) +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.6) +
-  labs(x="Household Income in 2019", y="Average Level of Anxiety")
+clean_data |>
+  ggplot(aes(x = language, fill = factor(anxiety))) +
+  geom_bar(position = "fill") +
+  scale_fill_manual(values = c("0" = "#f4bfb6", "1" = "#f09286"), 
+                    labels = c("No or Mild Anxiety", "Moderate Anxiety or more")) +
+  coord_flip() +
+  labs(fill = "Anxiety", x = "Language of the Survey", y = "Proportion") +
+  guides(fill = guide_legend(position = "bottom"))
+```
+
+```{r}
+#| label: fig-income_education_dep_anx
+#| fig-cap: Levels of Depression and Anxiety Among Pregnant Individuals Grouped By Household Income and Education
+#| echo: false
+#| warning: false
+#| message: false
+#| layout-ncol: 2
+#| fig-subcap: ["Income and Depression", "Education and Depression", "Income and Anxiety", "Education and Anxiety"]
+clean_data |> 
+  ggplot(aes(x=household_income,fill = factor(depression))) +
+  geom_bar(position = "fill") +
+  scale_fill_manual(values = c("0" = "lightcyan3", "1" = "cadetblue"), 
+                    labels = c("No or Mild Depression", "Moderate Depression or more"))+
+  coord_flip()+
+  labs(fill = "Depression", x = "Household Income in 2019", y = "Proportion") + 
+  guides(fill = guide_legend(position = "bottom"))
+
+clean_data |>
+  ggplot(aes(x = maternal_education, fill = factor(depression))) +
+  geom_bar(position = "fill") +
+  scale_fill_manual(values = c("0" = "lightcyan3", "1" = "cadetblue"), 
+                    labels = c("No or Mild Depression", "Moderate Depression or more")) +
+  coord_flip() +
+  labs(fill = "Depression", x = "Level of Education", y = "Proportion")+
+  guides(fill = guide_legend(position = "bottom"))
+
+clean_data |>
+  ggplot(aes(x = household_income, fill = factor(anxiety))) +
+  geom_bar(position = "fill") +
+  scale_fill_manual(values = c("0" = "#f4bfb6", "1" = "#f09286"), 
+                    labels = c("No or Mild Anxiety", "Moderately Anxiety or more")) +
+  coord_flip() +
+  labs(fill = "Anxiety", x = "Household Income in 2019", y = "Proportion")+
+  guides(fill = guide_legend(position = "bottom"))
+
+clean_data |>
+  ggplot(aes(x = maternal_education, fill = factor(anxiety))) +
+  geom_bar(position = "fill") +
+  scale_fill_manual(values = c("0" = "#f4bfb6", "1" = "#f09286"), 
+                    labels = c("No or Mild Anxiety", "Moderate Anxiety or more")) +
+  coord_flip() +
+  labs(fill = "Anxiety", x = "Level of Education", y = "Proportion")+
+  guides(fill = guide_legend(position = "bottom"))
 ```
@@
 ```
+
 ```{r}
@@
 ```{r}
-#| message: false
+#| label: fig-average_depression_and_anxiety
+#| fig-cap: Distribution of Levels of Depression and Anxiety by Education and Household Income
 #| echo: false
 #| warning: false
-#| label: fig-language_income_education
-#| fig-cap: Income and Education Groups Distribution Based on the Choice of Survey Language
+#| message: false
 #| layout-ncol: 2
-#| fig-subcap: ["Distribution of Income by Language", "Distribution of Education by Language"]
-clean_data |>
-  ggplot(aes(x = household_income, fill = language)) +
-  stat_count(position = "dodge") +
-  theme_minimal() +
-  coord_flip() +
-  scale_fill_brewer(palette = "Set1") + theme(legend.position = "bottom")
+#| fig-subcap: ["Distribution of Levels of Depression by Education", "Distribution of Levels of Depression by Household Income", "Distribution of Levels of Anxiety by Education", "Distribution of Levels of Anxiety by Household Income"]
+clean_data %>%
+  group_by(maternal_education) %>%
+  summarise(average_depression = mean(EPDS)) %>%
+  ggplot(aes(x = maternal_education, y = average_depression)) +
+  geom_bar(stat = "identity", fill = "cornflowerblue") + geom_text(aes(label = round(average_depression, 3)), vjust = -0.5) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.6) +
+  labs(x="Education", y="Average Level of Depression")
 
-clean_data |>
-  ggplot(aes(x = maternal_education, fill = language)) +
-  stat_count(position = "dodge") +
-  theme_minimal() +
-  coord_flip() +
-  scale_fill_brewer(palette = "Set1") + theme(legend.position = "bottom")
+clean_data %>%
+  group_by(household_income) %>%
+  summarise(average_depression = mean(EPDS)) %>%
+  ggplot(aes(x = household_income, y = average_depression)) + 
+  geom_text(aes(label = round(average_depression, 3)), vjust = -0.5) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.6)+
+  geom_bar(stat = "identity", fill = "cornflowerblue") + labs(x="Household Income in 2019", y="Average Level of Depression")
+
+clean_data %>%
+  group_by(maternal_education) %>%
+  summarise(average_anxiety = mean(PROMIS)) %>%
+  ggplot(aes(x = maternal_education, y = average_anxiety)) +
+  geom_bar(stat = "identity", fill = "salmon2") + geom_text(aes(label = round(average_anxiety, 3)), vjust = -0.5) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.6)+
+  labs(x="Education", y="Average Level of Anxiety")
+
+clean_data %>%
+  group_by(household_income) %>%
+  summarise(average_anxiety = mean(PROMIS)) %>%
+  ggplot(aes(x = household_income, y = average_anxiety)) + 
+  geom_bar(stat = "identity", fill = "salmon2") +
+  geom_text(aes(label = round(average_anxiety, 3)), vjust = -0.5) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ylim(0, 0.6) +
+  labs(x="Household Income in 2019", y="Average Level of Anxiety")
 ```

### 2024-03-29 scripts/02-clean.R
scripts modification

@@ -46,7 +46,9 @@ clean_data <- raw_data_complete |>  filter(year == '2020') |> filter(month != 'A
   select(age_group, maternal_age, income, maternal_education, birth_weight, birth_length,
          delivery_mode, language, threaten_life, threaten_baby_danger, threaten_baby_harm, 
          month, EPDS, PROMIS) |>
-  rename(household_income = income)
+  rename(household_income = income) |> 
+  mutate(depression = case_when(EPDS >= 0.4 ~ 1, EPDS < 0.4 ~ 0),
+         anxiety = case_when(PROMIS >= 0.464 ~ 1, PROMIS < 0.464 ~ 0))
 
 write.csv(
   x = clean_data,

### 2024-03-29 scripts/99-model.R

@@ -30,7 +30,7 @@ clean_data$month <- factor(clean_data$month, levels = month_levels)
 # Depression Model
 depression_model <-
   stan_glm(
-    EPDS ~ age_group + household_income + language + maternal_education + threaten_life + threaten_baby_danger,
+    depression ~ age_group + household_income + language + maternal_education + threaten_life + threaten_baby_danger,
     data = clean_data,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
@@ -47,7 +47,7 @@ saveRDS(
 # Anxiety Model
 anxiety_model <-
   stan_glm(
-    PROMIS ~ age_group + household_income + language + maternal_education + threaten_life + threaten_baby_danger,
+    anxiety ~ age_group + household_income + language + maternal_education + threaten_life + threaten_baby_danger,
     data = clean_data,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),

### 2024-03-31 other/datasheet/datasheet_template.qmd
finish the first draft of paper

---
title: "Datasheet for 'A dataset'"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to enable analysis of Australian politicians. We were unable to find a publicly available dataset in a structured format that had the biographical and political information on Australian politicians that was needed for modelling.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - TBD
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - TBD
4. *Any other comments?*
    - TBD

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- TBD
2. *How many instances are there in total (of each type, if appropriate)?*
	- TBD
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- TBD
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- TBD
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- TBD
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- TBD
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- TBD
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- TBD
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- TBD
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- TBD
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- TBD
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- TBD
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- TBD
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- TBD
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- TBD
16. *Any other comments?*
	- TBD

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- TBD
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- TBD
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- TBD
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- TBD
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- TBD
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- TBD
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- TBD
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- TBD
4. *Any other comments?*
	- TBD

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- TBD
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- TBD
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- TBD
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- TBD
7. *Any other comments?*
	- TBD

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- TBD
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- TBD
3. *Is there an erratum? If so, please provide a link or other access point.*
	- TBD
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- TBD
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- TBD
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- TBD
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- TBD
8. *Any other comments?*
	- TBD

\newpage

# References

### 2024-03-31 outputs/paper.qmd
finish the first draft of paper

@@
     abstract: "The outbreak of the Covid-19 pandemic was a substantial stressor, especially for pregnant individuals.
-This paper uses Bayesian model and logistic regression to estimate levels of prenatal depression and anxiety during this hardship.
-The analysis reveals that the perceived threats to oneself and the unborn child are positively correlated to the mental distress, especially among younger pregnant individuals with lower levels of education and from lower-income families."
-date: "27 March 2024"
+Since depression and anxiety beyond moderate levels are considered clinically significant, this paper uses Bayesian model and logistic regression to evaluate the probability of experiencing at least moderate levels of prenatal depression or anxiety during the pandemic.
+The analysis reveals that the perceived threats to oneself and the unborn child are strongly correlated to the mental distress, especially among younger pregnant individuals with lower levels of education and from lower-income families."
+date: "30 March 2024"
 date-format: "D MMMM YYYY"
@@
 # Introduction {#sec-introduction}
+While pregnancy and childbirth are often moments of joy, they can also bring about distress as mothers navigate physical changes, uncertainties and responsibilities.
+In the 12 months before childbirth, 12.5% of women in Canada reported that most days were very stressful, while 7.5% of women reported depressive or anxious symptoms [@canadaPregnancyWomenaposs].
+These figures are likely to increase during the COVID-19 pandemic due to the heightened uncertainty and panic.
+Maternal depression and anxiety during pregnancy are associated with adverse neurodevelopmental outcomes for the offspring, such as hyperactivity disorder, autism, poor cognitive or emotional developmental [@MPMHPbeta], and thus, prenatal mental health is crucial for both mothers and children.
+In light of this, I utilize data from the Open Science Framework [@osfPregnancyDuring] to explore factors contributing to maternal depression and anxiety during the pandemic, along with associated birth outcomes.
+This analysis could also shed light on mothers’ mental health during other objective hardships or the general effects of prenatal distress.
+
+Given that moderate depression and anxiety can impair daily functioning, this paper distinguishes between levels of distress and use logistic regression to estimate the likelihood of pregnant individuals experiencing at least moderate levels of depression and anxiety in Canada during the pandemic.
+Predictors include age, education, household income, language preference, and perceived threat to oneself and the unborn baby.
+The results indicate that young pregnant individuals with lower education and income, as well as a high perception of danger to both themselves and their unborn child, are more likely to experience at least moderate depression and anxiety.
+Furthermore, participants who completed the survey in French demonstrated a lower likelihood of getting distress compared to those who responded in English.
+While the dataset used in this paper provides valuable information on pregnant participants’ demographic information, basic birth outcomes, and mental health data, it does not include any information on newborns’ health.
+Consequently, this paper mainly focuses on the maternal well-being and its affects on mothers.
+
+This paper is constructed as follows: The data utilized for analysis is discussed in @sec-data.
+Following that, two models, one for depression and one for anxiety, is introduced in @sec-model.
+@sec-results then talks about the results derived from the models.
+Lastly, I will discuss the findings in a broader context and address limitations of this paper as well as the future explorations in @sec-discussion.
+This paper uses the programming language R [@r].
+The analysis, the model and the visualizations use the following packages: `rstanarm` [@rrstanarm], `modelsummary` [@rmodelsummary], `ggplot2` [@rggplot2], `dplyr` [@rdplyr], `readr` [@rreadr], `janitor` [@rjanitor], `arrow` [@rarrow], `here` [@rhere], `broom` [@rbroom], and `knitr` [@rknitr].
 ```{r}
@@
 # Data {#sec-data}
-## Data Source and Measurement
+## Data Overview and Measurement
 The prenatal mental health dataset from the Open Science Framework was designed to explore the relationship between exposure to objective hardship caused by the Covid -19 pandemic and psychological distress in pregnant individuals [@osfPregnancyDuring].
@@
 A portion of advertisements targeted geographic regions and sociodemographic groups with less representation to reduce underrepresentation [@nihPrenatalMental].
+The surveys were available in either English or French.
 Prenatal data, including age, 2019 household income in Canadian dollars, and education level, were collected as part of the national Pregnancy during the COVID-19 Pandemic (PdP) project using online questionnaires through REDCap.
 Birth data were also acquired using parents reports in REDCap [@nihPrenatalMental].
-Depression symptoms were self-assessed using the Edinburgh Postnatal Depression Scale (EPDS), consisting of 10 questions, each scored from 0 to 3, with possible scores ranging from 0 to 30.
-Anxiety symptoms were self-assessed by Patient-Reported Outcomes Measurement Information System (PROMIS), including 7 questions, each scored from 1 to 5, with possible total scores ranging from 7 to 35.
+Depression symptoms were self-assessed using the Edinburgh Postnatal Depression Scale (EPDS), consisting of 10 questions, each scored from 0 to 3, with possible scores ranging from 0 to 30 [@EDPSstanford].
+Anxiety symptoms were self-assessed by Patient-Reported Outcomes Measurement Information System (PROMIS), including 7 questions, each scored from 1 to 5, with possible total scores ranging from 7 to 35 [@psychiatry].
 Higher scores on both surveys indicate more severe depression or anxiety symptoms.
 In this paper, the EPDS scores are divided by 30, and the PROMIS scores are adjusted by subtracting 7 from the original scores and then dividing by 35, and hence, making the level of depression and anxiety between 0 and 1 to standardize the scale.
-The surveys were available in either English or French.
 
+Two new binary variables named `depression` and `anxiety` are created.
+Mothers scoring above 12, which represents 40% on the EPDS scale, are likely to be suffering from depression and should seek medical attention [@EDPSstanford].
+Consequently, the variable `depression` is assigned to a value of 1 if the EPDS score is greater than or equal to 0.4, and 0 otherwise.
+Similarly, the `PROMIS` score of at least 20 indicates at least moderate depression.
+Hence, the variable `anxiety` is assigned to 1 if the `PROMIS` is larger or equal to 0.464, and 0 if it is smaller than 0.464.
+
+This paper distinguishes between levels of mental health using the term ‘moderate’, as symptoms of moderate depression may become severe enough to influence daily life [@healthlineMildModerate].
+Additionally, though mild anxiety may not always meet clinical significance criteria, anxiety reaching a moderate level is generally considered significant [@betterhelpWhatDifferent].
+During the early stages of the pandemic, especially among pregnant individuals, there is a higher likelihood of experiencing depression or anxiety symptoms.
+Therefore, I will focus on the estimation of at least moderate depression/anxiety.
+In this analysis, “depression” refers to being at least moderately depressed, while “no depression” indicates not reaching the level of moderate depression; “anxiety” refers to being at least moderately anxious, while “no anxiety” indicates not reaching the level of moderate anxiety.
 Data are available for 10,772 participants.
 After excluding observations with missing values and setting the delivery month from May 2020 to December 2020, 3266 observations remain.
-Further details regarding data cleaning can be found in @sec-appendix.
+Further details regarding data cleaning can be found in @sec-appendix-cleaning.
 All participants in this study provided informed consent.
+Participants with higher education levels and higher household income exhibit lower levels of depression and anxiety.
 
+## Data Analysis
 ```{r}
@@
 ```
+[@fig-month] provides an overview of the levels of depression and anxiety spanning from May 2020 to December 2020.
+Overall, there is a decreasing trend observed in both scales with the occurrence of the pandemic.
+On average, the level of anxiety tends to be slightly higher than that of depression, with the majority experiencing depression levels below 0.5.
+However, depression exhibits more higher outliers, implying that some individuals experience significantly severe depression symptoms compared to the average.
+Mothers scoring above 12 or 13, which represents 40%, are likely to be suffering from depression and should seek medical attention [@EDPSstanford].
+Hence, the distributions in [@fig-month] indicate a considerable number of pregnant individuals may be experiencing distress, given that the means fall within the range of 0.3 to 0.5.
+
+The mean age of mothers at the time of delivery in Canada is approximately 31 [@statcanMeanMother].
+Since the dataset contains maternal ages ranging from 18.5 to 49, and 35 and above is categorized as advanced maternal age [@nihMaternalRisk], I have grouped the ages as follows: Under 27, 27 to 34, 35 to 39, and 40 and above.
+The distribution of pregnant individuals’ ages and household incomes is illustrated in [@fig-income_age], and the relationship between age and education is depicted in [@fig-education_age].
+Both graphs shows that a significant proportion of participants fall within the 27 to 34 age group, followed by those aged 35 to 39.
+Surprisingly, the number of participants under 27 is very low, almost comparable to those aged 40 and above.
+
+Regarding household income in 2019, a large proportion of respondents reported incomes ranging from 70,000 to 149,999 as indicated in [@fig-income_age].
+There is a notable presence of households with incomes exceeding 200,000 dollars, while very few households reported incomes below 40,000.
+For reference, the median after-tax income of Canadian households was 73,000 in 2020, and 11.1% of Canadians were in low-income group [@statcanIncomeCanada].
+[@fig-education_age] demonstrates that the majority of pregnant individuals in the survey hold undergraduate degrees, with some having completed college/trade school or earned Master’s degrees.
+Very few respondents reported having less than a high school diploma.
+According to Statistics Canada, in 2021, 26.3% of women+ aged 15 and above hold high school diploma, 21.2% had college, CEGEP or other non-university certificate or diploma.
+58.4% hold postsecondary certificate, while 5.9% completed Master’s Degree [@statcanHighestLevel].
 
@@
 
+\newpage
 ```{r}
@@
 ```
-
+Next, [@fig-age_language_dep_anx] presents the proportions of participants experiencing less than moderate levels of depression/anxiety and those experiencing at least moderate depression/anxiety, categorized by age and language.
+From [@fig-age_language_dep_anx] (a) and (c), younger participants under 27 are more prone to feeling depressed and anxious, with the difference in depression being more noticeable.
+[@fig-age_language_dep_anx] (b) and (d), reveal that participants who completed the survey in French have lower levels of distress.
+Given that French speakers are concentrated in Quebec, which was actually one of the provinces most affected by the pandemic in 2020 [@COVID19Tracker], I consider the language, potentially indicating region of residence, is a factor influencing the probability of significant distress among pregnant individuals during the Covid-19 pandemic
 ```{r}
@@
 
+[@fig-income_education_dep_anx] displays the proportions of participants with less than moderate levels of depression/anxiety and those with at least moderate depression/anxiety, categorized by household income in 2019 and education.
+This contrast is particularly clear in [@fig-income_education_dep_anx] (a) and (b), which visualize the level of depression.
+More details can be found in @sec-appendix-average.
 ```{r}
@@
 ```
+\newpage
+Three questions specific to fears of COVID-19 were asked in the survey.
+Answers were on a continuous scale from 0 (“not at all”) to 100 (“very much so”), with 50 representing “somewhat” [@nihPrenatalMental].
 
+1. How much do you think your life is in danger during the COVID-19 pandemic?
+2. How much do you think your unborn baby’s life is in danger during the COVID-19 pandemic?
+3. How much are you worried that exposure to the COVID-19 virus will harm your unborn baby?
+
+Given the similarity between questions 1 and 2, which respectively focus on self-perceived danger and danger perceived towards the unborn, I will analyze the responses to these questions, but the relationship between question 2 and 3 concerning the baby will be presented in @sec-appendix-questions.
+Scores are divided by 100, resulting in a range from 0 to 1.
+The relationship between the perceived level of danger, both towards self and the unborn, and prenatal mental health is visualized in [@fig-danger].
+
+[@fig-danger] (a) and (b) reveals that compared to perceived level of the unborn baby, a higher perceived level of danger to oneself tends to lend to a higher level of depression.
+From [@fig-danger] (c) and (d), it appears that the perceived level of danger to both oneself and the unborn baby seem to have similar impacts on the mental health of pregnant individuals.
+Again, the average level of distress among French speakers is lower than that among English speakers, but participants who speak French have a higher variation.
+
+Therefore, the predictors I choose to estimate the prenatal mental health are maternal age groups, household income, education, the perceived danger to oneself, and the perceived danger to the unborn baby.
 ```{r}
@@
 ```
-
+\newpage
 # Model {#sec-model}
@@
 ```
+## Model Set-up
+To estimate whether pregnant individuals will experience depression or anxiety beyond moderate levels, I will utilize logistic regression to create separate models for each psychological distress, using the `stan_glm()` function from the `rstanarm` package [@rrstanarm].
+The predictors are age group, education, household income, language, the perceived level of danger to oneself, and the perceived level of danger to the unborn baby.
+It is important to recognize that a model is not, and cannot be, a true representation of reality [@tellingstorieswithdata].
+The purpose of a model is to identify patterns in the data and make estimations, so models are not guaranteed to be 100% accurate.
+
+Logistic regression is well-suited when the outcome variable is binary.
+For these two models, the outcome variables `depression` and `anxiety` are assigned to 1 if participants exhibit at least moderate symptoms and 0 otherwise.
+The logistic regression equation for both models can be expressed as follows:
+$$
+\begin{aligned}
+y_i|\pi_i & \sim \mbox{Bern}(\pi_i) \\
+\mbox{logit}(\pi_i) &= \beta_0+ \beta_1 \times \mbox{Age}_i + \beta_2 \times \mbox{Education}_i  + \beta_3 \times \mbox{HouseholdIncome}_i + \beta_4 \times \mbox{Language}_i \\
+&+ \beta_5 \times \mbox{DangerToSelf}_i + \beta_6 \times \mbox{DangerToBaby}_i\\
+\beta_0 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_1 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_2 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_3 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_4 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_5 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_6 & \sim \mbox{Normal}(0, 2.5)
+\end{aligned}
+$$
+
+where:
+
+- $y_i$ is the dependent variable, representing the probability of experiencing at least moderate prenatal depression or anxiety
+-  $\beta_0$ is the intercept term, representing the value if all the predictors are zero. It has a prior distribution that is normal with a mean of 0 and a standard deviation of 2.5
+- $\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, $\beta_5$, $\beta_6$ are the coefficients corresponding to the predictor variables, each with a prior distribution that is normal with mean 0 and standard deviation 2.5
+
+## Model Justification
+For the depression model, I expect to see a positive relationship between experiencing at least moderate depression and age under 27, low levels of education (e.g.
+less than high school diploma), low income household in 2019 (e.g.
+less than 20,000), and preference of using English.
+This expectation is based on @sec-data.
+Moreover, at the onset of the pandemic, when everything was uncertain, it is reasonable to that participants with lower income would worry more about the pregnancy, because financial strain can contribute to limited access to resources such as healthcare.
+Younger individuals or those with lower levels of education may experience increased depression due to limited life experience or reduced access to information.
+Pregnant individuals who perceive a higher level of threat to themselves or their unborn baby may also experience heightened stress levels.
+
+For the anxiety model, I also expect a positive relationship between experiencing at least moderate depression and age under 27, low levels of education and income, a preference of speaking English, and high perception of danger to both self and unborn baby.
+This is based on @sec-data as well.
 
 # Results {#sec-results}
+## Depression
+The coefficients from the logistic regression model, presented in [@tbl-depression_model], reveal the impact of various predictors on whether a pregnant individual would experience at least moderate depression.
+Age groups 27 to 34, 35 to 39, and 40 and above all exhibit negative coefficients, suggesting a lower likelihood of depression in individuals aged 27 and above.
+The general trend in income is that a higher household income positively correlates with a lower probability of depression, but the highest income category, 200,000 and above, shows a slightly higher coefficient than the second higher income range of 175,000 to 199,999.
+Similarly, higher levels of education are associated with lower depression levels, with the highest level, the Doctoral Degree holders, having slightly larger coefficients than those with  the second highest degree, the Master’s degrees.
+Individuals who use French have a negative coefficient, implying a smaller likelihood of depression.
+The perceived threats to oneself and the unborn baby exhibit large positive coefficients, and the threat to life has a larger coefficient than the threat to the baby.
 ```{r}
@@
 ```
+A participant with Master’s Degree, a household income of 175,000 to 199,999, aged 27 to 34, who speaks French, and reports a low perceived threat is least likely to feel depressed.
+Conversely, a participant with less than a high school diploma, a household income between 20,000 to 39,999, aged under 27, who speaks English, and reports a high perception of danger is the most likely to experience depression.
 
+To provide intuitive insights into the model, I visualize the coefficients from [@tbl-depression_model] in [@fig-depression_interval] along with their 95% confident intervals.
+The age variables, the language, and threats variables have small confidence intervals.
+The intercept and the education variables display relatively large confidence intervals, suggesting lower precision in these estimates.
 ```{r}
@@
 
+## Anxiety
+The coefficients from the logistic regression model for anxiety, presented in Table [@tbl-anxiety_model], shows the influence of predictor variables on whether a pregnant individual would experience at least moderate anxiety.
+Age groups 27 to 34, 35 to 39, and 40 and above have negative coefficients from [@tbl-anxiety_model], indicating a reduced likelihood of anxiety in people aged 27 and above, particularly in the 35 to 39 age group.
+Higher household income is associated with a lower probability of anxiety, with the income range of 175,000 to 199,999 showing the lowest coefficient.
+While higher levels of education correspond to lower anxiety levels, the differences are less evident compared to the depression model.
+To be more specific, the highest degree Doctoral Degree has coefficient -0.695, which is between the coefficients of College/trade school and Undergraduate degree, with coefficients -0.642 and -0.705 respectively.
+Pregnant individuals that speak French has a negative coefficient which implies a lower likelihood of anxiety as well.
+The perceived threat to oneself and the unborn baby exhibit large positive coefficients and are larger than the coefficients in the depression model.
+The two values are similar, 1.505 and 1.555, respectively.
+
+A participant with Master’s Degree, a household income of 175,000 to 199,999, aged 35 to 39, who speaks French, and reports a low perceived threat is least likely to experience anxiety.
+Conversely, a participant with less than a high school diploma, a household income below 20,000, aged under 27, who speaks English, and reports a high perception of danger is the most likely to feel anxious.
 ```{r}
@@
 
+[@fig-anxiety_interval] visualizes the coefficients and 95% confidence intervals from [@tbl-anxiety_model].
+As with the depression model, the intercept and education variables display larger confidence intervals,  indicating lower precision in these estimates.
 ```{r}
@@
 ```
+# Discussion {#sec-discussion}
+## Depression and Anxiety
+As shown in @sec-results, lower household income is associated with higher levels of depression and anxiety, as financial strain can significantly impact individuals’ access to essential resources, including healthcare, particularly during uncertain times.
+Low education has been linked to a lack of a sense of control and resilience, and exposure to day-to-day stressors [@newsmedicalDoesEducation].
+These negative factors have been strongly correlated with the onset of depression and anxiety [@keytopics].
+In contrast, education is shown to be a clear indicators of life outcomes such as employment, income, and social status, and hence, a highly-educated participant is prone to having fewer symptoms [@keytopics].
 
-# Discussions {#sec-discussions}
+It is observed that pregnant individuals under 27 are more likely to experience depression and anxiety.
+This could be attributed to the fact that younger mothers have a higher likelihood of experiencing their first pregnancy compared to older mothers, and older mothers have a greater chance of having previously given birth.
+Young individuals often face various socioeconomic challenges, including lower income and limited time to complete their education, which can lead to lower educational attainment levels.
+These factors can contribute to a higher likelihood of experiencing depression and anxiety.
+
+When individuals perceive a threat, their bodies activate the stress response system, which is an activation of stress hormones that produce physiological changes [@nihPhysiologyStress].
+More than this, individuals who perceive threats may engage in rumination and a decreased sense of control, continuously dwelling on their fears and anxieties.
+Thus, a high perception of threat to oneself and baby may lead to physiological disorders.
+
+The differences in language preference is often related to the regions.
+Most people who prefer to speak French reside in Quebec.
+As Quebec being one of the provinces with the highest number of COVID-19 cases in 2020 [@COVID19Tracker], the reason behind French speakers having low level of depression and anxiety remains uncertain.
+It could potentially be attributed to lifestyle and cultural attitudes prevalent among Quebecers.
+Further research is needed to better understand the underlying factors.
+
+## Medical Treatment
+Only one-third of pregnant women with depression would consider taking anti-depression treatment [@NOWWUADCWWWD].
+Hence, it is critical to raise awareness of depression as a health issue and encourage individuals to seek professional help.
+Research has shown that pregnant individuals with untreated depression face significantly increased infant risks compared to those without depression, particularly in terms of preterm birth and small infant size.
+However, treating depression with antidepressants has also been  found associated with high risks of preterm birth and low birth weight [@NOWWUADCWWWD].
+In terms of these findings, non-pharmacological therapies may be more acceptable for pregnant individuals, even though conclusive evidence supporting these therapies is currently lacking [@NOWWUADCWWWD].
+
+## Birth Weight/Length Related to Prenatal Mental Health
+Study has showed that maternal distress may contribute to reduced birthweight, and complications related to preterm birth and low birth weight are leading causes of infant morbidity and mortality [@NOWWUADCWWWD].
+The dataset used in this paper, from [@osfPregnancyDuring], also demonstrate a reduced birth weight and birth length related to  poor mental health.
+As illustrated in [@fig-weight_length_delivery], there is a clear association between higher levels of depression or anxiety and lower birthweight and length, categorized by delivery mode, with Caesarean-section babies having a larger average length and weight compared to vaginally delivered babies.
+These findings, again, emphasize the importance of addressing maternal mental health during pregnancy to support fetal growth and development.
 ```{r}
@@
 ```
+## Bias, Weakness, and Limitation
+Participants in the study were pregnant individuals aged 17 years or older, with gestation periods of up to 35 weeks.
+However, during the COVID-19 outbreak, potential biases of the data may occur, such as stress among young mothers under 17 years old and participants with more than 35 weeks of gestation, potentially leading to more severe depression or anxiety.
+This sampling bias could result in an underestimation of mental distress levels among pregnant individuals.
+As demonstrated in @sec-results, younger mothers tend to experience higher levels of depression and anxiety, and they are less likely to be financially stable compared to adults.
+It is important to understand the mental health of mothers under 17 years old, which helps .
+
+In the survey, depression and anxiety levels were assessed through self-reported measures using the Edinburgh Postnatal Depression Scale and Promis Anxiety Scale, while birth length and weight were reported by parents.
+This reliance on self-reporting and parental reporting introduces biases and potential ethical implications that may limit the accuracy of data.
+
+One weakness of the dataset is its lack of information on other mental disorders besides depression and anxiety, like psychosis, and geographical data.
+Considering that regions with different population densities may experience varying levels of panic during the pandemic, the analysis could benefit from including relevant regional data for a better precision.
+Moreover, some surveys were collected through social media like Facebook and Instagram Ads.
+These platforms may not be perceived as formal or serious compared to surveys conducted by established survey companies, meaning that respondents may not have provided accurate or complete information, potentially affecting the reliability and validity of the data collected.
+
+While several rounds of surveys were conducted, the complete dataset containing additional information was not publicly accessible.
+Consequently, this paper only utilized one survey and could not analyze changes in mental distress over time or assess newborn health outcomes, representing a limitation of the study.
+
+## Next step
+Given the challenges posed by the COVID-19 pandemic and the escalation of conflicts and crises worldwide, understanding prenatal mental health amidst such objective hardships is useful and necessary.
+I recommend that researchers focus on investigating how distress during pregnancy impacts birth outcomes, including the health and developmental outcomes of newborns, as well as postnatal mental health.
+Future research can explore deeper into the subjective factors, such as the prolonged duration of staying at home and the attitudes of family members, in order to comprehensively assess their impacts on maternal well-being.
+
+The models generated from this analysis serve as a foundation for advancing discussions on resource allocation, clinical interventions and the role of familial support in aiding pregnant individuals.
+By identifying the multifaceted factors influencing maternal well-being and mental health outcomes, policymakers and healthcare providers can make informed decisions regarding resources allocation and targeted treatments to provide support for pregnant individuals through hardships.
+
 \newpage
 # Appendix {#sec-appendix}
+## Depression and Anxiety by Education and Income {#sec-appendix-average}
+[@fig-average_depression_and_anxiety] illustrates the detailed distribution of levels of depression and anxiety by education and household income.
+Generally, participants with higher education levels and higher household income exhibit lower levels of depression and anxiety.
+Focusing on the lowest education and lowest income groups, individuals with less than high school diploma are more easily to experience anxiety, with an average score of 0.529, compared to 0.499 among those with household incomes less than 20,000.
+Conversely, individuals with household incomes less than 20,000 tend to feel depressed, with an average score of 0.423, compared to 0.407 among those with less than a high school diploma.
 ```{r}
@@
 ```
-
+## Answers to Question 2 and 3 {#sec-appendix-questions}
+[@fig-danger_baby] displays the relationship between the answers to question 2 and question 3 in @sec-data.
 ```{r}
@@
 \newpage
+## Data cleaning {#sec-appendix-cleaning}
+After obtaining the original raw dataset, I select the following variables: `maternal_age`, `household_income`, `edinburgh_postnatal_depression_scale`, `promis_anxiety`, `delivery_date_converted_to_month_and_year`, `birth_length`, `birth_weight`, `maternal_education`, `delivery_mode`, `language`, `threaten_life`, `threaten_baby_danger`, `threaten_baby_harm`.
+I exclude those observations with missing values and create two new variables, `month` and `year`, derived from `delivery_date_converted_to_month_and_year`, representing the delivery month and year respectively.
+Additionally, I categorized maternal age into four groups: under 27, 27 to 34, 35 to 39, and 40 and above, creating a variable called `age_group`.
+I then convert `edinburgh_postnatal_depression_scale` and `promis_anxiety` to percentages ranging from 0 to 1, named `EPDS` and `PROMIS` as specified in @sec-data.
+I scale `threaten_life`, `threaten_baby_danger` and `threaten_baby_harm` by dividing each by 100 to ensure they range from 0 to 1.
+
+Subsequently, I filtered the dataset to retain only observations from the year 2020, excluding those with a delivery month of April, considering that the surveys were conducted from May 2020 onwards.
+Finally, two variables named `depression` and `anxiety` are created based on the criteria stated in @sec-data: `depression` is set to 1 if `EPDS` is greater than or equal to 0.4, and 0 otherwise; similarly, `anxiety` is assigned to 1 if `PROMIS` is not less than 0.464, and 0 if it is smaller than 0.464.
+
+\newpage
 # References

### 2024-03-31 other/datasheet/datasheet_template.qmd
add simulation

@@
 title: "Datasheet for 'A dataset'"
-subtitle: "My subtitle if needed"
+subtitle: "Analysis of Prenatal Mental Health During the Pandemic"
 author: 
-  - First author
-  - Another author
+  - Boxuan Yi
 thanks: "Code and data are available at: LINK."
-date: "`r Sys.time()`"
+date: "31 March 2024"
 date-format: "D MMMM YYYY"
@@
 1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
-    - The dataset was created to enable analysis of Australian politicians.
-We were unable to find a publicly available dataset in a structured format that had the biographical and political information on Australian politicians that was needed for modelling.
+    - The survey aimed to explore the associations among exposure to objective hardship caused by the pandemic, perceived stress and psychological distress in pregnant individuals, as well as developmental outcomes in their offspring.
+Although there is evidence that disasters increase symptoms of mental illness, research on the mental health consequences of pandemics is limited.
 2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
-    - TBD
+    - Authors are Gerald Giesbrecht, Catherine Lebel, and Lianne Tomfohr-Madsen from the University of Calgary.
 3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
-    - TBD
+    - This project is supported by funds from the Owerko Center at the Alberta Children’s Hospital Research Institute.
 4. *Any other comments?*
-    - TBD
+    - A few surveys and follow-ups are conducted, but only one is open-access.
 
@@
 Please provide a description.*
-	- TBD
+	- ID, demographic information, mental health, birth outcomes, and the use of language of the survey.
 2. *How many instances are there in total (of each type, if appropriate)?*
-	- TBD
+	- There are 16 variables.
+One is the ID, three about demographic information, two about depression and anxiety, six about birth outcomes, one is the use of language, and three are answers to the survey questions.
 3.
@@
 If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
-	- TBD
+	- Many surveys and follow-ups were conducted, but only one concluded dataset is open-access, which is the dataset I am using.
+Surveys were released at different times.
+Other than the variables in the dataset I use, there are some other variables such as the health information of the new born.
 4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
@@
 8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
-	- TBD
+	- No
 9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
-	- TBD
+	- No. The researchers excluded those observations with errors.
 10.
@@
 8. *Any other comments?*
-	- TBD
+	- No
 

### 2024-03-31 scripts/00-simulation.R

@@ -1,0 +1,26 @@
+# Purpose: Simulate the survey data about prenatal mental health
+# Author: Boxuan Yi
+# Email: boxuan.yi@mail.utoronto.ca
+# Date: 31 March 2024
+# Prerequisites: Download and get familiar with the survey dataset
+
+income = c("$200,000+", "$100,000 - $124,999", "$40,000 - $69,999", "$70,000 - $99,999", "$20,000 - $39,999", "$125,000 - $149,999", "$150,000 - $174,999","$175,000 - $199,999", "Less than $20,000")
+education = c('Masters degree','Undergraduate degree','College/trade school','Doctoral Degree','High school diploma','Less than high school diploma')
+month = c("May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")
+
+simulated_data <-
+  tibble("ID" = 1:500,
+         "maternal_age" = sample(17:50, 500, replace = TRUE),
+         "household_income" = sample(income, 500, replace = TRUE),
+         "education" = sample(education, 500, replace = TRUE),
+         "birthlength" = sample(40:65, 500, replace = TRUE),
+         "birthweight" = sample(2700:4800, 500, replace = TRUE),
+         "delivery_mode" = sample(c("c-section", "vaginally"), 500, replace = TRUE),
+         "language" = sample(c("English", "French"), 500, replace = TRUE),
+         "threaten_life" = sample(0:100, 500, replace = TRUE),
+         "threaten_baby_danger" = sample(0:100, 500, replace = TRUE),
+         "threaten_baby_harm" = sample(0:100, 500, replace = TRUE),
+         "month" = sample(month, 500, replace = TRUE),
+         "EPDS" = sample(0:30, 500, replace = TRUE),
+         "PROMIS" = sample(7:35, 500, replace = TRUE))
+simulated_data

### 2024-04-02 outputs/paper.qmd
modify the structure of the paper

@@
 The analysis reveals that the perceived threats to oneself and the unborn child are strongly correlated to the mental distress, especially among younger pregnant individuals with lower levels of education and from lower-income families."
-date: "30 March 2024"
+date: "2 April 2024"
 date-format: "D MMMM YYYY"
@@
 Predictors include age, education, household income, language preference, and perceived threat to oneself and the unborn baby.
+The estimand being explored is whether pregnant individuals, during the pandemic, would experience at least moderate levels of depression or anxiety, as self-reported using the Edinburgh Postnatal Depression Scale and the Patient-Reported Outcomes Measurement Information System Scales in the survey.
+
 The results indicate that young pregnant individuals with lower education and income, as well as a high perception of danger to both themselves and their unborn child, are more likely to experience at least moderate depression and anxiety.
@@
 ```
-[@fig-month] provides an overview of the levels of depression and anxiety spanning from May 2020 to December 2020.
+\newpage
+Firstly, [@fig-month] provides an overview of the levels of depression and anxiety spanning from May 2020 to December 2020.
 Overall, there is a decreasing trend observed in both scales with the occurrence of the pandemic.
@@
 Hence, the distributions in [@fig-month] indicate a considerable number of pregnant individuals may be experiencing distress, given that the means fall within the range of 0.3 to 0.5.
-
-The mean age of mothers at the time of delivery in Canada is approximately 31 [@statcanMeanMother].
-Since the dataset contains maternal ages ranging from 18.5 to 49, and 35 and above is categorized as advanced maternal age [@nihMaternalRisk], I have grouped the ages as follows: Under 27, 27 to 34, 35 to 39, and 40 and above.
-The distribution of pregnant individuals’ ages and household incomes is illustrated in [@fig-income_age], and the relationship between age and education is depicted in [@fig-education_age].
-Both graphs shows that a significant proportion of participants fall within the 27 to 34 age group, followed by those aged 35 to 39.
-Surprisingly, the number of participants under 27 is very low, almost comparable to those aged 40 and above.
-
-Regarding household income in 2019, a large proportion of respondents reported incomes ranging from 70,000 to 149,999 as indicated in [@fig-income_age].
-There is a notable presence of households with incomes exceeding 200,000 dollars, while very few households reported incomes below 40,000.
-For reference, the median after-tax income of Canadian households was 73,000 in 2020, and 11.1% of Canadians were in low-income group [@statcanIncomeCanada].
-[@fig-education_age] demonstrates that the majority of pregnant individuals in the survey hold undergraduate degrees, with some having completed college/trade school or earned Master’s degrees.
-Very few respondents reported having less than a high school diploma.
-According to Statistics Canada, in 2021, 26.3% of women+ aged 15 and above hold high school diploma, 21.2% had college, CEGEP or other non-university certificate or diploma.
-58.4% hold postsecondary certificate, while 5.9% completed Master’s Degree [@statcanHighestLevel].
 
@@
 
-\newpage
 ```{r}
@@
 ```
-Next, [@fig-age_language_dep_anx] presents the proportions of participants experiencing less than moderate levels of depression/anxiety and those experiencing at least moderate depression/anxiety, categorized by age and language.
+The mean age of mothers at the time of delivery in Canada is approximately 31 [@statcanMeanMother].
+Since the dataset contains maternal ages ranging from 18.5 to 49, and 35 and above is categorized as advanced maternal age [@nihMaternalRisk], I have grouped the ages as follows: Under 27, 27 to 34, 35 to 39, and 40 and above.
+The distribution of pregnant individuals’ ages and household incomes is illustrated in [@fig-income_age], and the relationship between age and education is depicted in [@fig-education_age].
+Both graphs shows that a significant proportion of participants fall within the 27 to 34 age group, followed by those aged 35 to 39.
+Surprisingly, the number of participants under 27 is very low, almost comparable to those aged 40 and above.
+
+Regarding household income in 2019, a large proportion of respondents reported incomes ranging from 70,000 to 149,999 as indicated in [@fig-income_age].
+There is a notable presence of households with incomes exceeding 200,000 dollars, while very few households reported incomes below 40,000.
+For reference, the median after-tax income of Canadian households was 73,000 in 2020, and 11.1% of Canadians were in low-income group [@statcanIncomeCanada].
+[@fig-education_age] demonstrates that the majority of pregnant individuals in the survey hold undergraduate degrees, with some having completed college/trade school or earned Master’s degrees.
+Very few respondents reported having less than a high school diploma.
+According to Statistics Canada, in 2021, 26.3% of women+ aged 15 and above hold high school diploma, 21.2% had college, CEGEP or other non-university certificate or diploma.
+58.4% hold postsecondary certificate, while 5.9% completed Master’s Degree [@statcanHighestLevel].
+
+\newpage
+Regarding age and language, [@fig-age_language_dep_anx] presents the proportions of participants experiencing less than moderate levels of depression/anxiety and those experiencing at least moderate depression/anxiety, categorized by age and language.
 From [@fig-age_language_dep_anx] (a) and (c), younger participants under 27 are more prone to feeling depressed and anxious, with the difference in depression being more noticeable.
@@
 #| label: fig-age_language_dep_anx
-#| fig-cap: Levels of Depression and Anxiety Among Pregnant Individuals Grouped By Age and Language
+#| fig-cap: Levels of Depression and Anxiety Among Pregnant Individuals Grouped By Age and Language Used to Complete the Survey
 #| echo: false
@@
   scale_fill_manual(values = c("0" = "lightcyan3", "1" = "cadetblue"), 
-                    labels = c("No or Mild Depression", "Moderate Depression or more"))+
+                    labels = c("None or \nMild Depression", "Moderate Depression \nor Higher"))+
   coord_flip()+
@@
   scale_fill_manual(values = c("0" = "lightcyan3", "1" = "cadetblue"), 
-                    labels = c("No or Mild Depression", "Moderate Depression or more")) +
+                    labels = c("None or \nMild Depression", "Moderate Depression \nor Higher")) +
   coord_flip() +
@@
   scale_fill_manual(values = c("0" = "#f4bfb6", "1" = "#f09286"), 
-                    labels = c("No or Mild Anxiety", "Moderately Anxiety or more")) +
+                    labels = c("None or \nMild Anxiety", "Moderate Anxiety \nor Higher")) +
   coord_flip() +
@@
   scale_fill_manual(values = c("0" = "#f4bfb6", "1" = "#f09286"), 
-                    labels = c("No or Mild Anxiety", "Moderate Anxiety or more")) +
+                    labels = c("None or \nMild Anxiety", "Moderate Anxiety \nor Higher")) +
   coord_flip() +

### 2024-04-02 outputs/paper.qmd
modify tests, llm and paper

@@
 Regarding age and language, [@fig-age_language_dep_anx] presents the proportions of participants experiencing less than moderate levels of depression/anxiety and those experiencing at least moderate depression/anxiety, categorized by age and language.
+Darker colors represent more severe lvel of depression or anxiety.
 From [@fig-age_language_dep_anx] (a) and (c), younger participants under 27 are more prone to feeling depressed and anxious, with the difference in depression being more noticeable.
@@
 [@fig-income_education_dep_anx] displays the proportions of participants with less than moderate levels of depression/anxiety and those with at least moderate depression/anxiety, categorized by household income in 2019 and education.
-This contrast is particularly clear in [@fig-income_education_dep_anx] (a) and (b), which visualize the level of depression.
+There is a correlation between higher education and higher household income with lower levels of both depression and anxiety.
+Conversely, individuals with lower education and income levels are more easily to experience at least moderate depression or anxiety.
+Comparing [@fig-income_education_dep_anx] (a) and (c), and [@fig-income_education_dep_anx] (b) and (d), we observed that a larger proportion of people experience anxiety compared to depression.
 More details can be found in @sec-appendix-average.
@@
 
-## Medical Treatment
+## Treatment
 Only one-third of pregnant women with depression would consider taking anti-depression treatment [@NOWWUADCWWWD].
 Hence, it is critical to raise awareness of depression as a health issue and encourage individuals to seek professional help.
+
 Research has shown that pregnant individuals with untreated depression face significantly increased infant risks compared to those without depression, particularly in terms of preterm birth and small infant size.
@@
 The dataset used in this paper, from [@osfPregnancyDuring], also demonstrate a reduced birth weight and birth length related to poor mental health.
-As illustrated in [@fig-weight_length_delivery], there is a clear association between higher levels of depression or anxiety and lower birthweight and length, categorized by delivery mode, with Caesarean-section babies having a larger average length and weight compared to vaginally delivered babies.
+As illustrated in [@fig-weight_length_delivery], there is a clear association between higher levels of depression or anxiety and lower birth weight and length, categorized by delivery mode, with Caesarean-section babies having a larger average length and weight compared to vaginally delivered babies.
 These findings, again, emphasize the importance of addressing maternal mental health during pregnancy to support fetal growth and development.
@@
 Given the challenges posed by the COVID-19 pandemic and the escalation of conflicts and crises worldwide, understanding prenatal mental health amidst such objective hardships is useful and necessary.
-I recommend that researchers focus on investigating how distress during pregnancy impacts birth outcomes, including the health and developmental outcomes of newborns, as well as postnatal mental health.
-Future research can explore deeper into the subjective factors, such as the prolonged duration of staying at home and the attitudes of family members, in order to comprehensively assess their impacts on maternal well-being.
+I recommend that researchers focus on investigating how distress during pregnancy impacts birth outcomes, including the health outcomes and the language development of newborns during social isolation, as well as the postnatal mental health.
+Future research can explore deeper into the subjective factors, such as the prolonged duration of staying at home and the family members' behaviors, in order to comprehensively assess their impacts on maternal well-being.
 

### 2024-04-02 scripts/00-simulation.R

@@ -4,6 +4,7 @@
 # Date: 31 March 2024
 # Prerequisites: Download and get familiar with the survey dataset
 
+# Simulation
 income = c("$200,000+", "$100,000 - $124,999", "$40,000 - $69,999", "$70,000 - $99,999", "$20,000 - $39,999", "$125,000 - $149,999", "$150,000 - $174,999","$175,000 - $199,999", "Less than $20,000")
 education = c('Masters degree','Undergraduate degree','College/trade school','Doctoral Degree','High school diploma','Less than high school diploma')
 month = c("May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")
@@ -12,15 +13,47 @@ simulated_data <-
   tibble("ID" = 1:500,
          "maternal_age" = sample(17:50, 500, replace = TRUE),
          "household_income" = sample(income, 500, replace = TRUE),
-         "education" = sample(education, 500, replace = TRUE),
-         "birthlength" = sample(40:65, 500, replace = TRUE),
-         "birthweight" = sample(2700:4800, 500, replace = TRUE),
+         "maternal_education" = sample(education, 500, replace = TRUE),
+         "birth_length" = sample(40:65, 500, replace = TRUE),
+         "birth_weight" = sample(2700:4800, 500, replace = TRUE),
          "delivery_mode" = sample(c("c-section", "vaginally"), 500, replace = TRUE),
          "language" = sample(c("English", "French"), 500, replace = TRUE),
          "threaten_life" = sample(0:100, 500, replace = TRUE),
          "threaten_baby_danger" = sample(0:100, 500, replace = TRUE),
          "threaten_baby_harm" = sample(0:100, 500, replace = TRUE),
          "month" = sample(month, 500, replace = TRUE),
-         "EPDS" = sample(0:30, 500, replace = TRUE),
-         "PROMIS" = sample(7:35, 500, replace = TRUE))
-simulated_data
+         "Edinburgh_Postnatal_Depression_Scale" = sample(0:30, 500, replace = TRUE),
+         "PROMIS_scale" = sample(7:35, 500, replace = TRUE))
+simulated_data <- simulated_data |> mutate("EPDS" = Edinburgh_Postnatal_Depression_Scale/30,
+                         "PROMIS" = (PROMIS_scale-7)/28)
+
+
+# Tests
+max(simulated_data$PROMIS) <= 1
+max(simulated_data$EPDS) <= 1
+max(simulated_data$threaten_life) <= 100
[…]

### 2024-04-02 outputs/paper.qmd
make the tests more clear and concise

@@
 ```
+
 ## Bias, Weakness, and Limitation

### 2024-04-02 scripts/03-test.R

@@ -21,18 +21,18 @@ min(clean_data$threaten_baby_danger) >= 0
 min(clean_data$threaten_baby_harm) >= 0
 
 # Groups and classes
-length(unique(clean_data$age_group)) <= 4
-length(unique(clean_data$household_income)) <= 9
-length(unique(clean_data$maternal_education)) <= 6
-is.numeric(clean_data$maternal_age)
-is.numeric(clean_data$birth_weight)
-is.numeric(clean_data$birth_length)
-is.numeric(clean_data$EPDS)
-is.numeric(clean_data$PROMIS)
+stopifnot(length(unique(clean_data$age_group)) <= 4,
+          length(unique(clean_data$household_income)) <= 9,
+          length(unique(clean_data$maternal_education)) <= 6,
+          is.numeric(clean_data$maternal_age),
+          is.numeric(clean_data$birth_weight),
+          is.numeric(clean_data$birth_length),
+          is.numeric(clean_data$EPDS),
+          is.numeric(clean_data$PROMIS))
 
-unique(clean_data$language) %in% c("English", "French")
-length(unique(clean_data$month)) <= 8
-length(unique(clean_data$delivery_mode)) == 2
+stopifnot(unique(clean_data$language) %in% c("English", "French"),
+          length(unique(clean_data$month)) <= 8,
+          length(unique(clean_data$delivery_mode)) == 2)
 
 class(clean_data$language) == "character"
 class(clean_data$delivery_mode) == "character"

### 2024-04-02 other/datasheet/datasheet.qmd
rename datasheet

--- 2024-03-31 other/datasheet/datasheet_template.qmd
@@
 ---
-title: "Datasheet for 'A dataset'"
-subtitle: "Analysis of Prenatal Mental Health During the Pandemic"
+title: "Datasheet for the Prenatal Mental Health Dataset"
 author: 
   - Boxuan Yi
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data are available at: https://github.com/Elaineyi1/Prenatal_Mental_Health"
 date: "31 March 2024"
 date-format: "D MMMM YYYY"
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
 format: pdf
 number-sections: true
-bibliography: references.bib
+bibliography: references2.bib
 ---
@@
 Extract of the questions from @gebru2021datasheets.
+
+The information provided below is sourced from:
+
+- https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10339202/ [@nihPrenatalMental] 
+- https://osf.io/ha5dp/ [@osfPregnancyDuring].
 
@@
 3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
-    - This project is supported by funds from the Owerko Center at the Alberta Children’s Hospital Research Institute.
+    -  This project was supported by funds from the Owerko Center at the Alberta Children’s Hospital Research Institute.
 4. *Any other comments?*
-    - A few surveys and follow-ups are conducted, but only one is open-access.
+    - No.
 
@@
 Please provide a description.*
-	- ID, demographic information, mental health, birth outcomes, and the use of language of the survey.
+	- The instances represent individuals, categorized into three types: ID, pregnant participants, and birth outcomes.
 2. *How many instances are there in total (of each type, if appropriate)?*
-	- There are 16 variables.
-One is the ID, three about demographic information, two about depression and anxiety, six about birth outcomes, one is the use of language, and three are answers to the survey questions.
+	- There are 3 instances and 16 variables in the dataset. One variable is the ID, nine variables are information about pregnant participants, and six contain birth outcomes.
 3.
@@
 If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
-	- Many surveys and follow-ups were conducted, but only one concluded dataset is open-access, which is the dataset I am using.
-Surveys were released at different times.
-Other than the variables in the dataset I use, there are some other variables such as the health information of the new born.
+	- Several surveys and follow-ups were conducted over time, but only one finalized dataset is available for open access, which is the dataset currently being utilized.
+Surveys were released at different stages.
+Apart from the variables included in the dataset being used, there are additional variables such as the health information of the new born.
 4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
-	- TBD
+	- The ID instance contains the unique ID assigned to each participant.
+The pregnant individuals instance contains demographic information and mental health data.
+The birth outcome instance include birth length, birthweight, and whether the new born was  admitted to the NICU.
 5. *Is there a label or target associated with each instance? If so, please provide a description.*
-	- TBD
+	- No.
 6.
@@
 This does not include intentionally removed information, but might include, for example, redacted text.*
-	- TBD
+	- Other than the variables included in the dataset being used, there are additional variables such as the health information of the new born, which are not open-access and hence missing.
 7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
-	- TBD
+	- Yes, the pregnant individuals were the mothers who gave birth to the newborns, and the birth outcomes are data specific to these newborns.
 8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
-	- No
+	- No.
 9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
@@
 Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
-	- TBD
+	- Yes, the dataset is self-contained. However, the data could be compared to pre-pandemic datasets to explore the impacts of the pandemic.
 11.
@@
 If so, please provide a description.*
-	- TBD
+	- No, and all the participants provided permission for the use of their data.
 12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
-	- TBD
+	- No.
 13.
@@
 If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
-	- TBD
+	- Yes.
+Demographic information including age, education, and household income was collected.
+These data could help understand how age and socioeconomic status may influence an individual’s level of depression and anxiety level during the pandemic.
 14.
@@
 If so, please describe how.*
-	- TBD
+	- No.
 15.
@@
 If so, please provide a description.*
-	- TBD
+	- This survey was designed to investigate prenatal mental health, so health data was included.
+The use of language, whether English or French, is also collected.
+This may indicate whether the participant is from Quebec or not.
 16. *Any other comments?*
-	- TBD
+	- No.
 
@@
 If so, please describe how.*
-	- TBD
+	- Data collection was conducted via online surveys. All the responses were self-reported or parent-reported.
 2.
@@
 How were these mechanisms or procedures validated?*
-	- TBD
+	- Data was collected through online questionnaires using REDCap.
+Participants were recruited via advertisements distributed through pregnancy organizations and care providers, social media, and via paid ads on Facebook and Instagram.
 3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
-	- TBD
+	- Participants were recruited via advertisements distributed through pregnancy organizations and care providers, social media, and via paid ads on Facebook and Instagram, with extra paid-ads focusing on underrepresentation groups.
 4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
-	- TBD
+	- Not mentioned.
 5.
@@
 If not, please describe the timeframe in which the data associated with the instances was created.*
-	- TBD
+	- April 2020 to April 2021. Yes.
 6.
@@
 If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
+	- No.
+All participants in this study provided informed consent and signed the electronic contract form.
+Research was approved by the University of Calgary conjoint health research ethics board (CHREB; REB20-0500) and conduced in accordance with the Declaration of Helsinki.
 7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
-	- TBD
+	- The dataset was available on the Open Science Framework at https://osf.io/ha5dp/
 8.
@@
 If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
-	- TBD
+	- All participants provided informed consent for research and signed the electronic consent form before starting the survey.
 9.
@@
 If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
-	- TBD
+	- All participants signed the electronic consent form before proceeding to the first questionnaire.
 10.
@@
 If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
-	- TBD
+	- Not mentioned.
 11.
@@
 If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
+	- Not mentioned.
 12. *Any other comments?*
-	- TBD
+	- No.
 
@@
 If not, you may skip the remaining questions in this section.*
-	- TBD
+	- Cases that seem ineligible, with insufficient information, duplicate entries, or nonsense information were removed.
+The researchers also looked at the time taken to complete the survey.
+All participants took at least 9 minutes to complete the surveys with sensible answers, suggesting that these were legitimate responses.
 2.
@@
 If so, please provide a link or other access point to the "raw" data.*
-	- TBD
+	- No
 3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
-	- TBD
+	- No
 4. *Any other comments?*
-	- TBD
+	- No
 
@@
 1. *Has the dataset been used for any tasks already? If so, please provide a description.*
-	- TBD
+	- Yes, the task aimed to understand the relationships among exposure to objective hardship caused by the pandemic, perceived stress in pregnant individuals, and developmental outcomes in their children.
 2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
-	- TBD
+	- No
 3. *What (other) tasks could the dataset be used for?*
-	- TBD
+	- To investigate how mothers’ mental health affect the babies.
 4.
@@
 Is there anything a dataset consumer could do to mitigate these risks or harms?*
-	- TBD
+	- No
 5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
-	- TBD
+	- No
 6. *Any other comments?*
-	- TBD
+	- No
 
@@
 If so, please provide a description.*
-	- TBD
+	- It permits unrestricted use, distribution, and reproduction in any medium, provided the original work, first published in JMIR Research Protocols, is properly cited.
 2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
-	- TBD
+	- This is open-access distributed under the terms of the Creative Commons Attribution License (https://creativecommons.org/licenses/by/4.0/)
 3. *When will the dataset be distributed?*
-	- TBD
+	- Not mentioned.
 4.
@@
 If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
-	- TBD
+	- Not mentioned.
 5.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
-	- TBD
+	- No
 6.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
-	- TBD
+	- No
 7. *Any other comments?*
-	- TBD
+	- No
 
@@
 1. *Who will be supporting/hosting/maintaining the dataset?*
-	- TBD
+	- The corresponding author, Gerald F Giesbrecht, PhD.
 2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
-	- TBD
+	-  The corresponding author, Gerald F Giesbrecht, PhD, provided the email ggiesbre@ucalgary.ca and phone number (+1)403-441-8469.
 3. *Is there an erratum? If so, please provide a link or other access point.*
-	- TBD
+	- No
 4.
@@
 If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
-	- TBD
+	- No, but the team claimed that 'the Pregnancy during the COVID-19 Pandemic (PdP) cohort study is ongoing, and we plan to release further data on pregnancy and child development as it is collected and cleaned'.
+However, I did not see any updates.
 5.
@@
 If so, please describe these limits and explain how they will be enforced.*
-	- TBD
+	- All the information gathered was anonymous, and participants provided consent for information use in the study.
 6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
-	- TBD
+	- The Pregnancy during the COVID-19 Pandemic cohort study was ongoing, and the team planed to release further data on pregnancy and child development as it was collected and cleaned.
 7.
@@
 If so, please provide a description.*
-	- TBD
+	- No, but the data could be compared to other datasets from the COVID-19 pandemic to establish generalizability, or to pre-pandemic datasets to determine the extent of changes during the COVID-19 pandemic.
 8. *Any other comments?*
-	- No
+	- No.
 

### 2024-04-07 outputs/paper.qmd
correct two grammar mistakes

@@
 Consequently, the variable `depression` is assigned to a value of 1 if the EPDS score is greater than or equal to 0.4, and 0 otherwise.
-Similarly, the `PROMIS` score of at least 20 indicates at least moderate depression.
-Hence, the variable `anxiety` is assigned to 1 if the `PROMIS` is larger or equal to 0.464, and 0 if it is smaller than 0.464.
+The PROMIS scale scoring more than 20 indicates at least moderate depression.
+Hence, the variable `anxiety` is assigned to 1 if the PROMIS score is larger or equal to 0.464, and 0 if it is smaller than 0.464.
 
@@
 This expectation is based on @sec-data.
-Moreover, at the onset of the pandemic, when everything was uncertain, it is reasonable to that participants with lower income would worry more about the pregnancy, because financial strain can contribute to limited access to resources such as healthcare.
+Moreover, at the onset of the pandemic, when everything was uncertain, it is reasonable that participants with lower income would worry more about the pregnancy, because financial strain can contribute to limited access to resources such as healthcare.
 Younger individuals or those with lower levels of education may experience increased depression due to limited life experience or reduced access to information.
@@
 Study has showed that maternal distress may contribute to reduced birthweight, and complications related to preterm birth and low birth weight are leading causes of infant morbidity and mortality [@NOWWUADCWWWD].
-The dataset used in this paper, from [@osfPregnancyDuring], also demonstrate a reduced birth weight and birth length related to poor mental health.
+The dataset used in this paper, from the Open Science Frame [@osfPregnancyDuring], also demonstrate a reduced birth weight and birth length related to poor mental health.
 As illustrated in [@fig-weight_length_delivery], there is a clear association between higher levels of depression or anxiety and lower birth weight and length, categorized by delivery mode, with Caesarean-section babies having a larger average length and weight compared to vaginally delivered babies.

### 2024-04-07 outputs/paper.qmd
correcting the grammar and add another conversation with chatgpt

--- 2024-04-02 outputs/paper.qmd
@@
 Since depression and anxiety beyond moderate levels are considered clinically significant, this paper uses Bayesian model and logistic regression to evaluate the probability of experiencing at least moderate levels of prenatal depression or anxiety during the pandemic.
-The analysis reveals that the perceived threats to oneself and the unborn child are strongly correlated to the mental distress, especially among younger pregnant individuals with lower levels of education and from lower-income families."
+The analysis indicates that younger age, low levels of education and household income, and a high perception of danger to both oneself and the unborn are associated with higher levels of depression and anxiety."
 date: "2 April 2024"
@@
 # Introduction {#sec-introduction}
-While pregnancy and childbirth are often moments of joy, they can also bring about distress as mothers navigate physical changes, uncertainties and responsibilities.
-In the 12 months before childbirth, 12.5% of women in Canada reported that most days were very stressful, while 7.5% of women reported depressive or anxious symptoms [@canadaPregnancyWomenaposs].
-These figures are likely to increase during the COVID-19 pandemic due to the heightened uncertainty and panic.
-Maternal depression and anxiety during pregnancy are associated with adverse neurodevelopmental outcomes for the offspring, such as hyperactivity disorder, autism, poor cognitive or emotional developmental [@MPMHPbeta], and thus, prenatal mental health is crucial for both mothers and children.
-In light of this, I utilize data from the Open Science Framework [@osfPregnancyDuring] to explore factors contributing to maternal depression and anxiety during the pandemic, along with associated birth outcomes.
-This analysis could also shed light on mothers’ mental health during other objective hardships or the general effects of prenatal distress.
+While pregnancy and childbirth are often moments of joy, they can also bring about distress as mothers navigate responsibilities and physical changes.
+In the 12 months before childbirth, 12.5% of women in Canada reported that most days were very stressful, and 7.5% reported depressive or anxious symptoms [@canadaPregnancyWomenaposs].
+These figures are likely to increase during the COVID-19 pandemic due to heightened uncertainty and panic.
+Maternal depression and anxiety during pregnancy are associated with adverse neurodevelopmental outcomes for the offspring, such as hyperactivity disorder, autism, poor cognitive or emotional development [@MPMHPbeta], which emphasizes the importance of prenatal mental health for both mothers and children.
+In light of this, I utilize data from the Open Science Framework to explore factors contributing to maternal depression and anxiety during the pandemic, along with associated birth outcomes.
+This analysis could also provide insight into mothers’ mental health during other objective hardships or the general effects of prenatal distress.
 
-Given that moderate depression and anxiety can impair daily functioning, this paper distinguishes between levels of distress and use logistic regression to estimate the likelihood of pregnant individuals experiencing at least moderate levels of depression and anxiety in Canada during the pandemic.
+Considering that moderate depression and anxiety can affect daily functioning, this paper distinguishes between levels of distress and use logistic regression to estimate the likelihood of pregnant individuals experiencing at least moderate levels of depression and anxiety in Canada during the pandemic.
 Predictors include age, education, household income, language preference, and perceived threat to oneself and the unborn baby.
@@
 
-The results indicate that young pregnant individuals with lower education and income, as well as a high perception of danger to both themselves and their unborn child, are more likely to experience at least moderate depression and anxiety.
-Furthermore, participants who completed the survey in French demonstrated a lower likelihood of getting distress compared to those who responded in English.
-While the dataset used in this paper provides valuable information on pregnant participants’ demographic information, basic birth outcomes, and mental health data, it does not include any information on newborns’ health.
-Consequently, this paper mainly focuses on the maternal well-being and its affects on mothers.
+The results indicate that young pregnant individuals with lower education and household income, as well as a high perception of danger to both themselves and their unborn child, are more likely to experience at least moderate depression and anxiety.
+Furthermore, participants who completed the survey in French demonstrate a lower probability of having mental distress compared to those who respond in English.
+Although the dataset used in this paper provides valuable information on pregnant participants’ demographic information, basic birth outcomes, and mental health data, it does not include any information on newborns’ health.
+Consequently, this paper mainly focuses on the maternal well-being and its effects on mothers.
 
 This paper is constructed as follows: The data utilized for analysis is discussed in @sec-data.
-Following that, two models, one for depression and one for anxiety, is introduced in @sec-model.
-@sec-results then talks about the results derived from the models.
+Following that, two models, one for depression and one for anxiety, are introduced in @sec-model.
+@sec-results then contains the results derived from the models.
 Lastly, I will discuss the findings in a broader context and address limitations of this paper as well as the future explorations in @sec-discussion.
@@
 The analysis, the model and the visualizations use the following packages: `rstanarm` [@rrstanarm], `modelsummary` [@rmodelsummary], `ggplot2` [@rggplot2], `dplyr` [@rdplyr], `readr` [@rreadr], `janitor` [@rjanitor], `arrow` [@rarrow], `here` [@rhere], `broom` [@rbroom], and `knitr` [@rknitr].
+
 ```{r}
@@
 ## Data Overview and Measurement
-The prenatal mental health dataset from the Open Science Framework was designed to explore the relationship between exposure to objective hardship caused by the Covid -19 pandemic and psychological distress in pregnant individuals [@osfPregnancyDuring].
-The  study population is pregnant individuals aged 17 or older, with gestation periods of no more than 35 weeks, residing in Canada.
-Data were collected from May 2020 to December 2021, containing demographic information, mental health data and basic birth outcomes [@nihPrenatalMental].
-For this paper, I will only use data collected from May 2020 to December 2020 to analyze the prenatal mental health during the onset of the outbreak, a period highly marked by panic and hardship.
+The prenatal mental health dataset sourced from the Open Science Framework was designed to investigate the relationship between exposure to objective hardship caused by the Covid -19 pandemic and psychological distress among pregnant individuals [@osfPregnancyDuring].
+The population is pregnant individuals aged 17 or older, with gestation periods not exceeding 35 weeks, residing in Canada.
+Data collection spanned from May 2020 to December 2021, containing demographic information, mental health data and basic birth outcomes [@nihPrenatalMental].
+For this paper, I will only use data collected with delivery month from May to December in 2020 to analyze the prenatal mental health during the initial phase of the outbreak, a period highly marked by panic and adversity.
 
-Participants were recruited through advertisements via pregnancy organizations, care providers, social media, and paid ads on Facebook and Instagram, with the chance to win a $500 gift card.
-A portion of advertisements targeted geographic regions and sociodemographic groups with less representation to reduce underrepresentation [@nihPrenatalMental].
-The surveys were available in either English or French.
-Prenatal data, including age, 2019 household income in Canadian dollars, and education level, were collected as part of the national Pregnancy during the COVID-19 Pandemic (PdP) project using online questionnaires through REDCap.
-Birth data were also acquired using parents reports in REDCap [@nihPrenatalMental].
-Depression symptoms were self-assessed using the Edinburgh Postnatal Depression Scale (EPDS), consisting of 10 questions, each scored from 0 to 3, with possible scores ranging from 0 to 30 [@EDPSstanford].
-Anxiety symptoms were self-assessed by Patient-Reported Outcomes Measurement Information System (PROMIS), including 7 questions, each scored from 1 to 5, with possible total scores ranging from 7 to 35 [@psychiatry].
-Higher scores on both surveys indicate more severe depression or anxiety symptoms.
-In this paper, the EPDS scores are divided by 30, and the PROMIS scores are adjusted by subtracting 7 from the original scores and then dividing by 35, and hence, making the level of depression and anxiety between 0 and 1 to standardize the scale.
+Participants were recruited through pregnancy organizations, care providers, social media, and paid ads on Facebook and Instagram, with the chance to win a 500-dollar gift card.
+Some of the advertisements were specifically targeted to geographic regions and sociodemographic groups with lower representation to reduce underrepresentation [@nihPrenatalMental].
+Surveys were available in either English or French.
+Prenatal data, including age, household income in Canadian dollars for the year 2019, and education level were collected as part of the national Pregnancy during the COVID-19 Pandemic (PdP) project using online questionnaires through REDCap.
+Birth data were also acquired using parents reports [@nihPrenatalMental].
 
-Two new binary variables named `depression` and `anxiety` are created.
-Mothers scoring above 12, which represents 40% on the EPDS scale, are likely to be suffering from depression and should seek medical attention [@EDPSstanford].
-Consequently, the variable `depression` is assigned to a value of 1 if the EPDS score is greater than or equal to 0.4, and 0 otherwise.
-Similarly, the `PROMIS` score of at least 20 indicates at least moderate depression.
-Hence, the variable `anxiety` is assigned to 1 if the `PROMIS` is larger or equal to 0.464, and 0 if it is smaller than 0.464.
+Depression symptoms were self-assessed using the Edinburgh Postnatal Depression Scale (EPDS), comprising 10 questions and each scored from 0 to 3, with possible scores ranging from 0 to 30 [@EDPSstanford].
+Anxiety symptoms were self-evaluated using the Patient-Reported Outcomes Measurement Information System (PROMIS), consisting of 7 questions scored from 1 to 5, with possible total scores ranging from 7 to 35 [@psychiatry].
+Higher scores on both scales indicate more severe depression or anxiety symptoms.
+To standardize the scale in this analysis, EPDS scores were divided by 30, and PROMIS scores were adjusted by subtracting 7 from the original scores and then dividing by 35, making the levels of depression and anxiety between 0 and 1.
 
-This paper distinguishes between levels of mental health using the term ‘moderate’, as symptoms of moderate depression may become severe enough to influence daily life [@healthlineMildModerate].
-Additionally, though mild anxiety may not always meet clinical significance criteria, anxiety reaching a moderate level is generally considered significant [@betterhelpWhatDifferent].
-During the early stages of the pandemic, especially among pregnant individuals, there is a higher likelihood of experiencing depression or anxiety symptoms.
-Therefore, I will focus on the estimation of at least moderate depression/anxiety.
-In this analysis, “depression” refers to being at least moderately depressed, while “no depression” indicates not reaching the level of moderate depression; “anxiety” refers to being at least moderately anxious, while “no anxiety” indicates not reaching the level of moderate anxiety.
-Data are available for 10,772 participants.
-After excluding observations with missing values and setting the delivery month from May 2020 to December 2020, 3266 observations remain.
-Further details regarding data cleaning can be found in @sec-appendix-cleaning.
+This paper distinguishes between levels of mental health using the term ‘moderate’, for the following reasons: 1.
+symptoms of moderate depression may become severe enough to impact daily life [@healthlineMildModerate]; 2.
+While mild anxiety may not always meet clinical significance criteria, anxiety reaching a moderate level is considered significant [@betterhelpWhatDifferent].
+Moreover, given the higher likelihood of experiencing depression or anxiety symptoms during the pandemic, particularly among pregnant individuals, the focus here will be on estimating at least moderate depression or anxiety.
+
+Two binary variables named `depression` and `anxiety` were created.
+Participants scoring above 12 on the EPDS scale, representing 40%, were likely to be experiencing depression and advised to seek medical attention [@EDPSstanford].
+Accordingly, the variable `depression` was assigned a value of 1 if the EPDS score was greater than or equal to 0.4, and 0 otherwise.
+For the PROMIS scale, a score exceeding 20 indicated at least moderate anxiety.
+Therefore, the variable `anxiety` was assigned to 1 if the PROMIS score was greater than or equal to 0.464, and 0 if it was less than 0.464, as 20 on a scale from 7 to 35 represents 46.4%.
+In this analysis, “depression” refers to being at least moderately depressed, while “no depression” indicates not reaching the level of moderate depression; “anxiety” refers to being at least moderately anxious, while “no anxiety” means not reaching the level of moderate anxiety.
+
+A total of 10,772 participants’ data were available.
+After excluding observations with missing values and setting the delivery month from May 2020 to December 2020, 3266 observations remained.
+Further details regarding data cleaning are provided in @sec-appendix-cleaning.
 All participants in this study provided informed consent.
-Participants with higher education levels and higher household income exhibit lower levels of depression and anxiety.
 
@@
 \newpage
-Firstly, [@fig-month] provides an overview of the levels of depression and anxiety spanning from May 2020 to December 2020.
-Overall, there is a decreasing trend observed in both scales with the occurrence of the pandemic.
-On average, the level of anxiety tends to be slightly higher than that of depression, with the majority experiencing depression levels below 0.5.
-However, depression exhibits more higher outliers, implying that some individuals experience significantly severe depression symptoms compared to the average.
-Mothers scoring above 12 or 13, which represents 40%, are likely to be suffering from depression and should seek medical attention [@EDPSstanford].
-Hence, the distributions in [@fig-month] indicate a considerable number of pregnant individuals may be experiencing distress, given that the means fall within the range of 0.3 to 0.5.
+Firstly, [@fig-month] provides an overview of the levels of depression and anxiety with delivery months spanning from May to December.
+Overall, a decreasing trend is observed in both scales with the onset of the pandemic.
+The level of anxiety tends to be slightly higher than that of depression, with the majority experiencing depression levels below 0.5.
+However, depression exhibits more outliers towards higher values, implying that some individuals experienced significantly severe depression symptoms.
+Mothers scoring above 0.4 in depression and 0.464 in anxiety are experiencing at least moderate distress.
+The distributions in [@fig-month] indicate a considerable number of pregnant individuals may be experiencing distress, given that the means fall within the range of 0.3 to 0.5.
 
@@
 ```
-The mean age of mothers at the time of delivery in Canada is approximately 31 [@statcanMeanMother].
-Since the dataset contains maternal ages ranging from 18.5 to 49, and 35 and above is categorized as advanced maternal age [@nihMaternalRisk], I have grouped the ages as follows: Under 27, 27 to 34, 35 to 39, and 40 and above.
+In Canada, the mean age of mothers at the time of delivery is approximately 31 [@statcanMeanMother].
+The dataset contains maternal ages ranging from 18.5 to 49, and because 35 and above is categorized as advanced maternal age [@nihMaternalRisk], I have grouped the ages as follows: Under 27, 27 to 34, 35 to 39, and 40 and above.
 The distribution of pregnant individuals’ ages and household incomes is illustrated in [@fig-income_age], and the relationship between age and education is depicted in [@fig-education_age].
-Both graphs shows that a significant proportion of participants fall within the 27 to 34 age group, followed by those aged 35 to 39.
+Both graphs show that a significant proportion of participants fall within the 27 to 34 age group, followed by those aged 35 to 39.
 Surprisingly, the number of participants under 27 is very low, almost comparable to those aged 40 and above.
 
-Regarding household income in 2019, a large proportion of respondents reported incomes ranging from 70,000 to 149,999 as indicated in [@fig-income_age].
-There is a notable presence of households with incomes exceeding 200,000 dollars, while very few households reported incomes below 40,000.
-For reference, the median after-tax income of Canadian households was 73,000 in 2020, and 11.1% of Canadians were in low-income group [@statcanIncomeCanada].
+A significant proportion of respondents reported incomes ranging from 70,000 to 149,999, as indicated in [@fig-income_age].
+There is a notable presence of households with incomes exceeding 200,000, while very few households reported incomes below 40,000.
+For reference, the median after-tax income of Canadian households was 73,000 in 2020, and 11.1% of Canadians were in the low-income group [@statcanIncomeCanada].
 [@fig-education_age] demonstrates that the majority of pregnant individuals in the survey hold undergraduate degrees, with some having completed college/trade school or earned Master’s degrees.
 Very few respondents reported having less than a high school diploma.
-According to Statistics Canada, in 2021, 26.3% of women+ aged 15 and above hold high school diploma, 21.2% had college, CEGEP or other non-university certificate or diploma.
-58.4% hold postsecondary certificate, while 5.9% completed Master’s Degree [@statcanHighestLevel].
+According to Statistics Canada, in 2021, 26.3% of women+ aged 15 and above hold a high school diploma, 21.2% had college, CEGEP, or other non-university certificate or diploma.
+58.4% hold a postsecondary certificate, while 5.9% completed a Master’s Degree [@statcanHighestLevel].
 
@@
 Regarding age and language, [@fig-age_language_dep_anx] presents the proportions of participants experiencing less than moderate levels of depression/anxiety and those experiencing at least moderate depression/anxiety, categorized by age and language.
-Darker colors represent more severe lvel of depression or anxiety.
+Darker colours represent a more severe level of depression or anxiety.
 From [@fig-age_language_dep_anx] (a) and (c), younger participants under 27 are more prone to feeling depressed and anxious, with the difference in depression being more noticeable.
-[@fig-age_language_dep_anx] (b) and (d), reveal that participants who completed the survey in French have lower levels of distress.
-Given that French speakers are concentrated in Quebec, which was actually one of the provinces most affected by the pandemic in 2020 [@COVID19Tracker], I consider the language, potentially indicating region of residence, is a factor influencing the probability of significant distress among pregnant individuals during the Covid-19 pandemic
+[@fig-age_language_dep_anx] (b) and (d) reveal that participants who completed the survey in French have lower levels of distress.
+Given that French speakers are concentrated in Quebec, which was one of the provinces most affected by the pandemic in 2020 [@COVID19Tracker], I consider the language, potentially indicating the region of residence, as a factor influencing the probability of significant distress among pregnant individuals during the Covid-19 pandemic.
+
 ```{r}
@@
 There is a correlation between higher education and higher household income with lower levels of both depression and anxiety.
-Conversely, individuals with lower education and income levels are more easily to experience at least moderate depression or anxiety.
-Comparing [@fig-income_education_dep_anx] (a) and (c), and [@fig-income_education_dep_anx] (b) and (d), we observed that a larger proportion of people experience anxiety compared to depression.
+Conversely, individuals with lower education and income levels are more likely to experience at least moderate depression or anxiety.
+Comparing [@fig-income_education_dep_anx] (a) and (c), and [@fig-income_education_dep_anx] (b) and (d), we observe that a larger proportion of people experience anxiety compared to depression.
 More details can be found in @sec-appendix-average.
@@
 \newpage
-Three questions specific to fears of COVID-19 were asked in the survey.
-Answers were on a continuous scale from 0 (“not at all”) to 100 (“very much so”), with 50 representing “somewhat” [@nihPrenatalMental].
+Three questions specific to fears of COVID-19 were asked in the survey. Answers were on a numeric scale from 0 (not at all) to 100 (very much so), with 50 representing ‘somewhat’ [@nihPrenatalMental].
 
@@
 
-Given the similarity between questions 1 and 2, which respectively focus on self-perceived danger and danger perceived towards the unborn, I will analyze the responses to these questions, but the relationship between question 2 and 3 concerning the baby will be presented in @sec-appendix-questions.
+Given the similarity between questions 1 and 2, which respectively focus on self-perceived danger and danger perceived towards the unborn, I will analyze the responses to these two questions, but the relationship between question 2 and 3 concerning the baby will be presented in @sec-appendix-questions.
 Scores are divided by 100, resulting in a range from 0 to 1.
@@
 
-[@fig-danger] (a) and (b) reveals that compared to perceived level of the unborn baby, a higher perceived level of danger to oneself tends to lend to a higher level of depression.
-From [@fig-danger] (c) and (d), it appears that the perceived level of danger to both oneself and the unborn baby seem to have similar impacts on the mental health of pregnant individuals.
+[@fig-danger] (a) and (b) reveal that compared to the perceived level of the unborn baby, a higher perceived level of danger to oneself tends to lead to a higher level of depression.
+From [@fig-danger] (c) and (d), it appears that the perceived level of danger to both oneself and the unborn baby have similar impacts on anxiety.
 Again, the average level of distress among French speakers is lower than that among English speakers, but participants who speak French have a higher variation.
@@
 Therefore, the predictors I choose to estimate the prenatal mental health are maternal age groups, household income, education, the perceived danger to oneself, and the perceived danger to the unborn baby.
+
 ```{r}
@@
 It is important to recognize that a model is not, and cannot be, a true representation of reality [@tellingstorieswithdata].
-The purpose of a model is to identify patterns in the data and make estimations, so models are not guaranteed to be 100% accurate.
+The purpose of a model is to find patterns in the data and make estimations, so models are not guaranteed to be 100% accurate.
 
@@
 The logistic regression equation for both models can be expressed as follows:
+
 $$
@@
 ## Model Justification
-For the depression model, I expect to see a positive relationship between experiencing at least moderate depression and age under 27, low levels of education (e.g.
+For the depression model, I expect to see a positive relationship between experiencing depression and age under 27, low levels of education (e.g.
 less than high school diploma), low income household in 2019 (e.g.
-less than 20,000), and preference of using English.
+less than 20,000), and a preference of using English.
 This expectation is based on @sec-data.
-Moreover, at the onset of the pandemic, when everything was uncertain, it is reasonable to that participants with lower income would worry more about the pregnancy, because financial strain can contribute to limited access to resources such as healthcare.
+Moreover, at the onset of the pandemic, when everything was uncertain, it is reasonable that participants with lower income would worry more about the pregnancy, because financial strain may lead to limited access to resources such as healthcare.
 Younger individuals or those with lower levels of education may experience increased depression due to limited life experience or reduced access to information.
@@
 
-For the anxiety model, I also expect a positive relationship between experiencing at least moderate depression and age under 27, low levels of education and income, a preference of speaking English, and high perception of danger to both self and unborn baby.
+For the anxiety model, I also expect a positive relationship between experiencing anxiety and age under 27, low levels of education and income, a preference of speaking English, and high perception of danger to both self and unborn baby.
 This is based on @sec-data as well.
@@
 ## Depression
-The coefficients from the logistic regression model, presented in [@tbl-depression_model], reveal the impact of various predictors on whether a pregnant individual would experience at least moderate depression.
+The coefficients from the logistic regression model, presented in [@tbl-depression_model], show the impact of various predictors on whether a pregnant individual would experience at least moderate depression.
 Age groups 27 to 34, 35 to 39, and 40 and above all exhibit negative coefficients, suggesting a lower likelihood of depression in individuals aged 27 and above.
-The general trend in income is that a higher household income positively correlates with a lower probability of depression, but the highest income category, 200,000 and above, shows a slightly higher coefficient than the second higher income range of 175,000 to 199,999.
+The general trend in income is that a higher household income positively correlates with a lower probability of depression, 200,000 and above, shows a slightly higher coefficient than the second higher income range of 175,000 to 199,999.
 Similarly, higher levels of education are associated with lower depression levels, with the highest level, the Doctoral Degree holders, having slightly larger coefficients than those with the second highest degree, the Master’s degrees.
 Individuals who use French have a negative coefficient, implying a smaller likelihood of depression.
-The perceived threats to oneself and the unborn baby exhibit large positive coefficients, and the threat to life has a larger coefficient than the threat to the baby.
+The perceived threats to oneself and the unborn baby exhibit large positive coefficients.
+The threat to life has a larger coefficient than the threat to the baby, meaning that feeling the threat to life can lead to more severe depression.
+
 ```{r}
@@
 The intercept and the education variables display relatively large confidence intervals, suggesting lower precision in these estimates.
+
 ```{r}
@@
 ## Anxiety
-The coefficients from the logistic regression model for anxiety, presented in Table [@tbl-anxiety_model], shows the influence of predictor variables on whether a pregnant individual would experience at least moderate anxiety.
-Age groups 27 to 34, 35 to 39, and 40 and above have negative coefficients from [@tbl-anxiety_model], indicating a reduced likelihood of anxiety in people aged 27 and above, particularly in the 35 to 39 age group.
+The coefficients from the logistic regression model for anxiety, presented in [@tbl-anxiety_model], indicate the influence of predictor variables on whether a pregnant individual would experience at least moderate anxiety.
+Age groups 27 to 34, 35 to 39, and 40 and above have negative coefficients from [@tbl-anxiety_model], suggesting a smaller likelihood of anxiety in individuals aged 27 and above, particularly in the 35 to 39 age group.
 Higher household income is associated with a lower probability of anxiety, with the income range of 175,000 to 199,999 showing the lowest coefficient.
-While higher levels of education correspond to lower anxiety levels, the differences are less evident compared to the depression model.
-To be more specific, the highest degree Doctoral Degree has coefficient -0.695, which is between the coefficients of College/trade school and Undergraduate degree, with coefficients -0.642 and -0.705 respectively.
-Pregnant individuals that speak French has a negative coefficient which implies a lower likelihood of anxiety as well.
-The perceived threat to oneself and the unborn baby exhibit large positive coefficients and are larger than the coefficients in the depression model.
+While higher levels of education correspond to lower anxiety, the differences are less evident compared to the depression model.
+To be more specific, the highest degree Doctoral Degree has coefficient of -0.695, which falls between the coefficients of College/trade school and Undergraduate degree, with coefficients of -0.642 and -0.705 respectively.
+Pregnant individuals who speak French have a negative coefficient, which implies a lower likelihood of anxiety as well.
+The perceived threat to oneself and the unborn baby have large positive coefficients, and they are larger than the coefficients in the depression model.
 The two values are similar, 1.505 and 1.555, respectively.
+The perception of threat to the baby leads to more severe anxiety symptoms.
 
@@
 Conversely, a participant with less than a high school diploma, a household income below 20,000, aged under 27, who speaks English, and reports a high perception of danger is the most likely to feel anxious.
+
 ```{r}
@@
 As with the depression model, the intercept and education variables display larger confidence intervals,  indicating lower precision in these estimates.
+
 ```{r}
@@
 ## Depression and Anxiety
-As shown in @sec-results, lower household income is associated with higher levels of depression and anxiety, as financial strain can significantly impact individuals’ access to essential resources, including healthcare, particularly during uncertain times.
-Low education has been linked to a lack of a sense of control and resilience, and exposure to day-to-day stressors [@newsmedicalDoesEducation].
+As demonstrated in @sec-results, lower household income is associated with higher levels of depression and anxiety, as financial strain can greatly impact individuals’ access to essential resources, including healthcare, particularly during uncertain times.
+Low education has been linked to a lack of a sense of control and resilience, as well as exposure to day-to-day stressors [@newsmedicalDoesEducation].
 These negative factors have been strongly correlated with the onset of depression and anxiety [@keytopics].
-In contrast, education is shown to be a clear indicators of life outcomes such as employment, income, and social status, and hence, a highly-educated participant is prone to having fewer symptoms [@keytopics].
+In contrast, education is shown to be a clear indicator of life outcomes such as employment, income, and social status, and hence, highly-educated participants are prone to having fewer symptoms [@keytopics].
 
@@
 
-When individuals perceive a threat, their bodies activate the stress response system, which is an activation of stress hormones that produce physiological changes [@nihPhysiologyStress].
-More than this, individuals who perceive threats may engage in rumination and a decreased sense of control, continuously dwelling on their fears and anxieties.
-Thus, a high perception of threat to oneself and baby may lead to physiological disorders.
+When individuals perceive a threat, their bodies activate the stress response system, which involves an activation of stress hormones that produce physiological changes [@nihPhysiologyStress].
+More than this, individuals who perceive threats may engage in rumination and experience a decreased sense of control, continuously dwelling on their fears and anxieties.
+Thus, a high perception of threat to oneself and the baby may lead to physiological disorders.
 
-The differences in language preference is often related to the regions.
+The differences in language preference are often related to the regions.
 Most people who prefer to speak French reside in Quebec.
-As Quebec being one of the provinces with the highest number of COVID-19 cases in 2020 [@COVID19Tracker], the reason behind French speakers having low level of depression and anxiety remains uncertain.
+Despite Quebec being one of the provinces with the highest number of COVID-19 cases in 2020 [@COVID19Tracker], the reason behind French speakers having a low level of depression and anxiety remains uncertain.
 It could potentially be attributed to lifestyle and cultural attitudes prevalent among Quebecers.
@@
 Research has shown that pregnant individuals with untreated depression face significantly increased infant risks compared to those without depression, particularly in terms of preterm birth and small infant size.
-However, treating depression with antidepressants has also been found associated with high risks of preterm birth and low birth weight [@NOWWUADCWWWD].
-In terms of these findings, non-pharmacological therapies may be more acceptable for pregnant individuals, even though conclusive evidence supporting these therapies is currently lacking [@NOWWUADCWWWD].
+However, treating depression with antidepressants has also been found to be associated with higher risks of preterm birth and low birth weight [@NOWWUADCWWWD].
+Considering these findings, non-pharmacological therapies may be more acceptable for pregnant individuals, and they usually provide a more personalized treatment for individuals.
 
 ## Birth Weight/Length Related to Prenatal Mental Health
-Study has showed that maternal distress may contribute to reduced birthweight, and complications related to preterm birth and low birth weight are leading causes of infant morbidity and mortality [@NOWWUADCWWWD].
-The dataset used in this paper, from [@osfPregnancyDuring], also demonstrate a reduced birth weight and birth length related to poor mental health.
+Studies have shown that maternal distress may contribute to reduced birthweight, and complications related to preterm birth and low birth weight are leading causes of infant morbidity and mortality [@NOWWUADCWWWD].
+The dataset used in this paper, from the Open Science Framework [@osfPregnancyDuring], also demonstrates a correlation between poor mental health and reduced birth weight and birth length.
 As illustrated in [@fig-weight_length_delivery], there is a clear association between higher levels of depression or anxiety and lower birth weight and length, categorized by delivery mode, with Caesarean-section babies having a larger average length and weight compared to vaginally delivered babies.
-These findings, again, emphasize the importance of addressing maternal mental health during pregnancy to support fetal growth and development.
+These findings, once again, emphasize the importance of addressing maternal mental health during pregnancy to support fetal growth and development.
+
 ```{r}
@@
 Participants in the study were pregnant individuals aged 17 years or older, with gestation periods of up to 35 weeks.
-However, during the COVID-19 outbreak, potential biases of the data may occur, such as stress among young mothers under 17 years old and participants with more than 35 weeks of gestation, potentially leading to more severe depression or anxiety.
-This sampling bias could result in an underestimation of mental distress levels among pregnant individuals.
+However, during the COVID-19 outbreak, potential biases in the data may occur, such as more severe depression or anxiety symptoms among young mothers under 17 years old and participants with more than 35 weeks of gestation.
+This sampling bias could result in an underestimation of mental distress levels.
 As demonstrated in @sec-results, younger mothers tend to experience higher levels of depression and anxiety, and they are less likely to be financially stable compared to adults.
-It is important to understand the mental health of mothers under 17 years old, which helps .
+Consequently, it is also important to understand the mental health of mothers under 17 years old.
 
@@
 
+Based on [@fig-income_age] and [@fig-education_age], it appears that the proportion of individuals with lower income and education levels is smaller than expected compared to the actual education [@statcanHighestLevel] and income distribution [@statcanIncomeCanada] in Canada.
+Individuals from lower socioeconomic backgrounds may face limited access to networks and healthcare services, which could hinder their participation in surveys or online platforms, thus skewing the results.
+They may also be too stressed to look at their phone or engage with social media, and then miss the survey advertisements.
+It is also possible that self-reporting inaccuracies about low income and education have influenced the data.
+Regardless of the reason, the underrepresentation has resulted in bias.
+
 One weakness of the dataset is its lack of information on other mental disorders besides depression and anxiety, like psychosis, and geographical data.
-Considering that regions with different population densities may experience varying levels of panic during the pandemic, the analysis could benefit from including relevant regional data for a better precision.
-Moreover, some surveys were collected through social media like Facebook and Instagram Ads.
+Considering that regions with different population densities may experience varying levels of panic during the pandemic, the analysis could benefit from including relevant regional data for better precision.
+Moreover, some surveys were collected through social media platforms like Facebook and Instagram Ads.
 These platforms may not be perceived as formal or serious compared to surveys conducted by established survey companies, meaning that respondents may not have provided accurate or complete information, potentially affecting the reliability and validity of the data collected.
@@
 While several rounds of surveys were conducted, the complete dataset containing additional information was not publicly accessible.
-Consequently, this paper only utilized one survey and could not analyze changes in mental distress over time or assess newborn health outcomes, representing a limitation of the study.
+Consequently, this paper only utilized one survey and could not analyze changes in mental distress over time or assess newborn health outcomes, and this is a limitation of the study.
 
 ## Next step
-Given the challenges posed by the COVID-19 pandemic and the escalation of conflicts and crises worldwide, understanding prenatal mental health amidst such objective hardships is useful and necessary.
-I recommend that researchers focus on investigating how distress during pregnancy impacts birth outcomes, including the health outcomes and the language development of newborns during social isolation, as well as the postnatal mental health.
-Future research can explore deeper into the subjective factors, such as the prolonged duration of staying at home and the family members' behaviors, in order to comprehensively assess their impacts on maternal well-being.
+Given the challenges posed by the COVID-19 pandemic and the escalation of conflicts and crises worldwide, understanding prenatal mental health amidst such objective hardships is both useful and necessary.
+I recommend that researchers focus on investigating how distress during pregnancy impacts birth outcomes, including the health outcomes and language development of newborns during social isolation, as well as postnatal mental health.
+Future research can explore deeper into subjective factors, such as the prolonged duration of staying at home and the behaviours of family members, in order to fully assess maternal well-being.
 
-The models generated from this analysis serve as a foundation for advancing discussions on resource allocation, clinical interventions and the role of familial support in aiding pregnant individuals.
-By identifying the multifaceted factors influencing maternal well-being and mental health outcomes, policymakers and healthcare providers can make informed decisions regarding resources allocation and targeted treatments to provide support for pregnant individuals through hardships.
+The models generated from this analysis serve as a foundation for advancing discussions on resource allocation, clinical interventions, and the role of familial support in caring pregnant individuals.
+By identifying the multifaceted factors influencing maternal well-being and mental health outcomes, policymakers and healthcare providers can make informed decisions regarding resource allocation and targeted treatments to provide support for pregnant individuals through hardships.
 

### 2024-04-08 outputs/paper.qmd
finalize the paper

@@
     abstract: "The outbreak of the Covid-19 pandemic was a substantial stressor, especially for pregnant individuals.
-Since depression and anxiety beyond moderate levels are considered clinically significant, this paper uses Bayesian model and logistic regression to evaluate the probability of experiencing at least moderate levels of prenatal depression or anxiety during the pandemic.
+Since depression and anxiety beyond moderate levels are considered clinically significant, this paper uses the Bayesian model and logistic regression to evaluate the probability of experiencing at least moderate levels of prenatal depression or anxiety during the pandemic.
 The analysis indicates that younger age, low levels of education and household income, and a high perception of danger to both oneself and the unborn are associated with higher levels of depression and anxiety."
-date: "2 April 2024"
+date: "8 April 2024"
 date-format: "D MMMM YYYY"
@@
 In the 12 months before childbirth, 12.5% of women in Canada reported that most days were very stressful, and 7.5% reported depressive or anxious symptoms [@canadaPregnancyWomenaposs].
-These figures are likely to increase during the COVID-19 pandemic due to heightened uncertainty and panic.
+These figures were likely to increase during the COVID-19 pandemic due to heightened uncertainty and panic.
 Maternal depression and anxiety during pregnancy are associated with adverse neurodevelopmental outcomes for the offspring, such as hyperactivity disorder, autism, poor cognitive or emotional development [@MPMHPbeta], which emphasizes the importance of prenatal mental health for both mothers and children.
@@
 
-Considering that moderate depression and anxiety can affect daily functioning, this paper distinguishes between levels of distress and use logistic regression to estimate the likelihood of pregnant individuals experiencing at least moderate levels of depression and anxiety in Canada during the pandemic.
+Considering that moderate depression and anxiety can affect daily functioning, this paper distinguishes between levels of distress and uses logistic regression to estimate the likelihood of pregnant individuals experiencing at least moderate levels of depression and anxiety in Canada during the pandemic.
 Predictors include age, education, household income, language preference, and perceived threat to oneself and the unborn baby.
@@
 The results indicate that young pregnant individuals with lower education and household income, as well as a high perception of danger to both themselves and their unborn child, are more likely to experience at least moderate depression and anxiety.
-Furthermore, participants who completed the survey in French demonstrate a lower probability of having mental distress compared to those who respond in English.
+Furthermore, participants who completed the survey in French demonstrate a lower probability of having mental distress compared to those who responded in English.
 Although the dataset used in this paper provides valuable information on pregnant participants’ demographic information, basic birth outcomes, and mental health data, it does not include any information on newborns’ health.
-Consequently, this paper mainly focuses on the maternal well-being and its effects on mothers.
+Consequently, this paper mainly focuses on maternal well-being and its effects on mothers.
 
-This paper is constructed as follows: The data utilized for analysis is discussed in @sec-data.
+This paper is constructed as follows: The data utilized for analysis is included in @sec-data.
 Following that, two models, one for depression and one for anxiety, are introduced in @sec-model.
-@sec-results then contains the results derived from the models.
+@sec-results then talks about the results derived from the models.
 Lastly, I will discuss the findings in a broader context and address limitations of this paper as well as the future explorations in @sec-discussion.
@@
 ## Data Overview and Measurement
-The prenatal mental health dataset sourced from the Open Science Framework was designed to investigate the relationship between exposure to objective hardship caused by the Covid -19 pandemic and psychological distress among pregnant individuals [@osfPregnancyDuring].
+The prenatal mental health dataset sourced from the Open Science Framework was designed to investigate the relationship between exposure to objective hardship caused by the Covid-19 pandemic and psychological distress among pregnant individuals [@osfPregnancyDuring].
 The population is pregnant individuals aged 17 or older, with gestation periods not exceeding 35 weeks, residing in Canada.
-Data collection spanned from May 2020 to December 2021, containing demographic information, mental health data and basic birth outcomes [@nihPrenatalMental].
+Data collection spanned from May 2020 to December 2021, containing individual-level demographic information, mental health data and basic birth outcomes [@nihPrenatalMental].
 For this paper, I will only use data collected with delivery month from May to December in 2020 to analyze the prenatal mental health during the initial phase of the outbreak, a period highly marked by panic and adversity.
-
 Participants were recruited through pregnancy organizations, care providers, social media, and paid ads on Facebook and Instagram, with the chance to win a 500-dollar gift card.
@@
 
-Depression symptoms were self-assessed using the Edinburgh Postnatal Depression Scale (EPDS), comprising 10 questions and each scored from 0 to 3, with possible scores ranging from 0 to 30 [@EDPSstanford].
+Depression symptoms were self-assessed using the Edinburgh Postnatal Depression Scale (EPDS), comprising ten questions and each scored from 0 to 3, with possible scores ranging from 0 to 30 [@EDPSstanford].
 Anxiety symptoms were self-evaluated using the Patient-Reported Outcomes Measurement Information System (PROMIS), consisting of 7 questions scored from 1 to 5, with possible total scores ranging from 7 to 35 [@psychiatry].
@@
 The distributions in [@fig-month] indicate a considerable number of pregnant individuals may be experiencing distress, given that the means fall within the range of 0.3 to 0.5.
-
 ```{r}
@@
 ```
-
+\newpage
 ```{r}
@@
 ```
-In Canada, the mean age of mothers at the time of delivery is approximately 31 [@statcanMeanMother].
+
+In Canada, the mean age of mothers at delivery is approximately 31 [@statcanMeanMother].
 The dataset contains maternal ages ranging from 18.5 to 49, and because 35 and above is categorized as advanced maternal age [@nihMaternalRisk], I have grouped the ages as follows: Under 27, 27 to 34, 35 to 39, and 40 and above.
@@
 A significant proportion of respondents reported incomes ranging from 70,000 to 149,999, as indicated in [@fig-income_age].
-There is a notable presence of households with incomes exceeding 200,000, while very few households reported incomes below 40,000.
+There is a notable presence of households with incomes exceeding 200,000, while very few reported incomes below 40,000.
 For reference, the median after-tax income of Canadian households was 73,000 in 2020, and 11.1% of Canadians were in the low-income group [@statcanIncomeCanada].
@@
 Very few respondents reported having less than a high school diploma.
-According to Statistics Canada, in 2021, 26.3% of women+ aged 15 and above hold a high school diploma, 21.2% had college, CEGEP, or other non-university certificate or diploma.
-58.4% hold a postsecondary certificate, while 5.9% completed a Master’s Degree [@statcanHighestLevel].
-
+According to Statistics Canada, in 2021, 26.3% of women+ aged 15 and above held a high school diploma, 21.2% had a college, CEGEP, or other non-university certificate or diploma.
+58.4% held a postsecondary certificate, while 5.9% completed a Master’s Degree [@statcanHighestLevel].
 \newpage
@@
 Given that French speakers are concentrated in Quebec, which was one of the provinces most affected by the pandemic in 2020 [@COVID19Tracker], I consider the language, potentially indicating the region of residence, as a factor influencing the probability of significant distress among pregnant individuals during the Covid-19 pandemic.
-
 ```{r}
@@
 
-[@fig-income_education_dep_anx] displays the proportions of participants with less than moderate levels of depression/anxiety and those with at least moderate depression/anxiety, categorized by household income in 2019 and education.
-There is a correlation between higher education and higher household income with lower levels of both depression and anxiety.
-Conversely, individuals with lower education and income levels are more likely to experience at least moderate depression or anxiety.
-Comparing [@fig-income_education_dep_anx] (a) and (c), and [@fig-income_education_dep_anx] (b) and (d), we observe that a larger proportion of people experience anxiety compared to depression.
-More details can be found in @sec-appendix-average.
 ```{r}
@@
 ```
+[@fig-income_education_dep_anx] displays the proportions of participants with less than moderate levels of depression/anxiety and those with at least moderate depression/anxiety, categorized by household income in 2019 and education.
+There is a correlation between higher education and higher household income with lower levels of both depression and anxiety.
+Conversely, individuals with lower education and income levels are more likely to experience at least moderate depression or anxiety.
+Comparing [@fig-income_education_dep_anx] (a) and (c), and [@fig-income_education_dep_anx] (b) and (d), we observe that a larger proportion of people experience anxiety compared to depression.
+More details can be found in @sec-appendix-average.
+
 \newpage
@@
 
-Given the similarity between questions 1 and 2, which respectively focus on self-perceived danger and danger perceived towards the unborn, I will analyze the responses to these two questions, but the relationship between question 2 and 3 concerning the baby will be presented in @sec-appendix-questions.
+Given the similarity between questions 1 and 2, which respectively focus on self-perceived danger and danger perceived towards the unborn, I will analyze the responses to these two questions, while the relationship between questions 2 and 3 concerning the baby will be presented in @sec-appendix-questions.
 Scores are divided by 100, resulting in a range from 0 to 1.
@@
 
-Therefore, the predictors I choose to estimate the prenatal mental health are maternal age groups, household income, education, the perceived danger to oneself, and the perceived danger to the unborn baby.
+Therefore, the predictors I choose to estimate prenatal mental health are maternal age groups, household income, education, the perceived danger to oneself, and the perceived danger to the unborn baby.
 
@@
 
-For the anxiety model, I also expect a positive relationship between experiencing anxiety and age under 27, low levels of education and income, a preference of speaking English, and high perception of danger to both self and unborn baby.
+For the anxiety model, I also expect a positive relationship between experiencing anxiety and age under 27, low levels of education and income, a preference of speaking English, and a high perception of danger to both self and unborn baby.
 This is based on @sec-data as well.
@@
 The general trend in income is that a higher household income positively correlates with a lower probability of depression, 200,000 and above, shows a slightly higher coefficient than the second higher income range of 175,000 to 199,999.
-Similarly, higher levels of education are associated with lower depression levels, with the highest level, the Doctoral Degree holders, having slightly larger coefficients than those with the second highest degree, the Master’s degrees.
+Similarly, higher levels of education are associated with lower depression levels, with the highest level, the Doctoral Degree holders, having slightly larger coefficients than those with the second highest degree, the Master’s degree.
 Individuals who use French have a negative coefficient, implying a smaller likelihood of depression.
@@
 ```
-A participant with Master’s Degree, a household income of 175,000 to 199,999, aged 27 to 34, who speaks French, and reports a low perceived threat is least likely to feel depressed.
+
+A participant with a Master’s Degree, a household income of 175,000 to 199,999, aged 27 to 34, who speaks French, and reports a low perceived threat is least likely to feel depressed.
 Conversely, a participant with less than a high school diploma, a household income between 20,000 to 39,999, aged under 27, who speaks English, and reports a high perception of danger is the most likely to experience depression.
 
-To provide intuitive insights into the model, I visualize the coefficients from [@tbl-depression_model] in [@fig-depression_interval] along with their 95% confident intervals.
-The age variables, the language, and threats variables have small confidence intervals.
+To provide intuitive insights into the model, I visualize the coefficients from [@tbl-depression_model] in [@fig-depression_interval] along with their 95% confidence intervals.
+The age, the language, and threats variables have small confidence intervals.
 The intercept and the education variables display relatively large confidence intervals, suggesting lower precision in these estimates.
@@
 While higher levels of education correspond to lower anxiety, the differences are less evident compared to the depression model.
-To be more specific, the highest degree Doctoral Degree has coefficient of -0.695, which falls between the coefficients of College/trade school and Undergraduate degree, with coefficients of -0.642 and -0.705 respectively.
+To be more specific, the highest degree Doctoral Degree has a coefficient of -0.695, which falls between the coefficients of College/trade school and Undergraduate degree, with coefficients of -0.642 and -0.705 respectively.
 Pregnant individuals who speak French have a negative coefficient, which implies a lower likelihood of anxiety as well.
-The perceived threat to oneself and the unborn baby have large positive coefficients, and they are larger than the coefficients in the depression model.
+The perceived threats to oneself and the unborn baby have large positive coefficients, and they are larger than the coefficients in the depression model.
 The two values are similar, 1.505 and 1.555, respectively.
 The perception of threat to the baby leads to more severe anxiety symptoms.
-
-A participant with Master’s Degree, a household income of 175,000 to 199,999, aged 35 to 39, who speaks French, and reports a low perceived threat is least likely to experience anxiety.
-Conversely, a participant with less than a high school diploma, a household income below 20,000, aged under 27, who speaks English, and reports a high perception of danger is the most likely to feel anxious.
 
@@
 ```
+
+A participant with a Master’s Degree, a household income of 175,000 to 199,999, aged 35 to 39, who speaks French, and reports a low perceived threat is least likely to experience anxiety.
+Conversely, a participant with less than a high school diploma, a household income below 20,000, aged under 27, who speaks English, and who reports a high perception of danger is the most likely to feel anxious.
 

### 2024-04-15 outputs/paper.qmd
finalize the paper based on the review

@@
 Predictors include age, education, household income, language preference, and perceived threat to oneself and the unborn baby.
+Although the dataset used in this paper provides valuable information on pregnant participants’ demographic information, basic birth outcomes, and mental health data, it does not include any information on newborns’ health.
+Consequently, this paper mainly focuses on maternal well-being and its effects on mothers.
+
 The estimand being explored is whether pregnant individuals, during the pandemic, would experience at least moderate levels of depression or anxiety, as self-reported using the Edinburgh Postnatal Depression Scale and the Patient-Reported Outcomes Measurement Information System Scales in the survey.
-
 The results indicate that young pregnant individuals with lower education and household income, as well as a high perception of danger to both themselves and their unborn child, are more likely to experience at least moderate depression and anxiety.
 Furthermore, participants who completed the survey in French demonstrate a lower probability of having mental distress compared to those who responded in English.
-Although the dataset used in this paper provides valuable information on pregnant participants’ demographic information, basic birth outcomes, and mental health data, it does not include any information on newborns’ health.
-Consequently, this paper mainly focuses on maternal well-being and its effects on mothers.
 
@@
 The predictors are age group, education, household income, language, the perceived level of danger to oneself, and the perceived level of danger to the unborn baby.
+The parameters in the models are considered random variables, each with its own associated probability distribution.
 It is important to recognize that a model is not, and cannot be, a true representation of reality [@tellingstorieswithdata].
@@
 
-- $y_i$ is the dependent variable, representing the probability of experiencing at least moderate prenatal depression or anxiety
+- $y_i$ is the dependent variable, representing the probability of a pregnant individual experiencing at least moderate level of depression or anxiety
 -  $\beta_0$ is the intercept term, representing the value if all the predictors are zero. It has a prior distribution that is normal with a mean of 0 and a standard deviation of 2.5
@@
 less than high school diploma), low income household in 2019 (e.g.
-less than 20,000), and a preference of using English.
+less than $20,000), and a preference of using English.
 This expectation is based on @sec-data.
@@
 Younger individuals or those with lower levels of education may experience increased depression due to limited life experience or reduced access to information.
-Pregnant individuals who perceive a higher level of threat to themselves or their unborn baby may also experience heightened stress levels.
+Pregnant individuals who perceive a higher level of threat to themselves or their unborn baby are also expected to experience heightened stress levels, because our bodies activate the stress response system when perceiving threats, leading to the release of stress hormones [@nihPhysiologyStress].
 
 For the anxiety model, I also expect a positive relationship between experiencing anxiety and age under 27, low levels of education and income, a preference of speaking English, and a high perception of danger to both self and unborn baby.
-This is based on @sec-data as well.
+This is based on @sec-data and the reasons listed in the above paragraph.
 

### 2024-04-17 outputs/paper.qmd
final :P

@@
     abstract: "The outbreak of the Covid-19 pandemic was a substantial stressor, especially for pregnant individuals.
-Since depression and anxiety beyond moderate levels are considered clinically significant, this paper uses the Bayesian model and logistic regression to evaluate the probability of experiencing at least moderate levels of prenatal depression or anxiety during the pandemic.
+Since depression and anxiety beyond moderate levels are considered clinically significant, this paper uses logistic regression to evaluate the probability of experiencing at least moderate levels of prenatal depression or anxiety in Canada during the pandemic.
 The analysis indicates that younger age, low levels of education and household income, and a high perception of danger to both oneself and the unborn are associated with higher levels of depression and anxiety."
-date: "8 April 2024"
+date: "17 April 2024"
 date-format: "D MMMM YYYY"
@@
 #| label: fig-month
-#| fig-cap: Distribution of Levels of Depression and Anxiety by Month
+#| fig-cap: The Distribution of Levels of Depression and Anxiety by Month
 #| echo: false
@@
 58.4% held a postsecondary certificate, while 5.9% completed a Master’s Degree [@statcanHighestLevel].
+The participants have a relatively higher income and education distribution than average.
 \newpage

### 2024-04-18 outputs/paper.qmd
final

@@
 #| label: tbl-depression_model
-#| tbl-cap: Estimating Prenatal Depression given Age, Household Income, Education Level and Choice of Language
+#| tbl-cap: Estimating Prenatal Depression given Age, Household Income, Education Level, Perception of Threats, and Choice of Language
 modelsummary(
@@
 #| label: tbl-anxiety_model
-#| tbl-cap: Estimating Prenatal Anxiety given Age Group, Household Income, Education Level and Choice of Language.
+#| tbl-cap: Estimating Prenatal Anxiety given Age Group, Household Income, Education Level, Perception of Treats, and Choice of Language
 modelsummary(

### 2024-04-18 scripts/00-simulation.R

@@ -28,6 +28,7 @@ simulated_data <- simulated_data |> mutate("EPDS" = Edinburgh_Postnatal_Depressi
                          "PROMIS" = (PROMIS_scale-7)/28)
 
 
+
 # Tests
 max(simulated_data$PROMIS) <= 1
 max(simulated_data$EPDS) <= 1

### 2024-04-18 scripts/03-test.R

@@ -7,6 +7,7 @@ clean_data <- read_csv(
   show_col_types = FALSE
 )
 
+
 # Maximum and minimum of scores
 max(clean_data$PROMIS) <= 1
 max(clean_data$EPDS) <= 1
@@ -19,6 +20,7 @@ min(clean_data$EPDS) >= 0
 min(clean_data$threaten_life) >= 0
 min(clean_data$threaten_baby_danger) >= 0
 min(clean_data$threaten_baby_harm) >= 0
+
 
 # Groups and classes
 stopifnot(length(unique(clean_data$age_group)) <= 4,

### 2024-04-18 outputs/paper.qmd
improve measurement

@@
 Maternal depression and anxiety during pregnancy are associated with adverse neurodevelopmental outcomes for the offspring, such as hyperactivity disorder, autism, poor cognitive or emotional development [@MPMHPbeta], which emphasizes the importance of prenatal mental health for both mothers and children.
-In light of this, I utilize data from the Open Science Framework to explore factors contributing to maternal depression and anxiety during the pandemic, along with associated birth outcomes.
+In light of this, I utilize data collected from the University of Calgary and Alberta Children's Hospital Research Institute to explore factors contributing to maternal depression and anxiety during the pandemic, along with associated birth outcomes.
 This analysis could also provide insight into mothers’ mental health during other objective hardships or the general effects of prenatal distress.
@@
 ## Data Overview and Measurement
-The prenatal mental health dataset sourced from the Open Science Framework was designed to investigate the relationship between exposure to objective hardship caused by the Covid-19 pandemic and psychological distress among pregnant individuals [@osfPregnancyDuring].
+The prenatal mental health dataset created by the University of Calgary and Alberta Children's Hospital Research Institute was designed to investigate the relationship between exposure to objective hardship caused by the Covid-19 pandemic and psychological distress among pregnant individuals [@osfPregnancyDuring].
 The population is pregnant individuals aged 17 or older, with gestation periods not exceeding 35 weeks, residing in Canada.
 Data collection spanned from May 2020 to December 2021, containing individual-level demographic information, mental health data and basic birth outcomes [@nihPrenatalMental].
+The dataset was downloaded from Open Science Framework.
 For this paper, I will only use data collected with delivery month from May to December in 2020 to analyze the prenatal mental health during the initial phase of the outbreak, a period highly marked by panic and adversity.
+
 Participants were recruited through pregnancy organizations, care providers, social media, and paid ads on Facebook and Instagram, with the chance to win a 500-dollar gift card.
@@
 Studies have shown that maternal distress may contribute to reduced birthweight, and complications related to preterm birth and low birth weight are leading causes of infant morbidity and mortality [@NOWWUADCWWWD].
-The dataset used in this paper, from the Open Science Framework [@osfPregnancyDuring], also demonstrates a correlation between poor mental health and reduced birth weight and birth length.
+The dataset used in this paper also demonstrates a correlation between poor mental health and reduced birth weight and birth length.
 As illustrated in [@fig-weight_length_delivery], there is a clear association between higher levels of depression or anxiety and lower birth weight and length, categorized by delivery mode, with Caesarean-section babies having a larger average length and weight compared to vaginally delivered babies.

### 2024-04-19 scripts/03-test.R
create a preamble and correct a typo

@@ -1,3 +1,9 @@
+#### Preamble ####
+# Purpose: Test the data
+# Author: Boxuan Yi
+# Email: boxuan.yi@mail.utoronto.ca
+# Date: 24 March 2024
+# Prerequisites: None
 library(arrow)
 library(readr)
 library(here)
